```{r, echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_termination_decay_correlation/'), 
                      dev='pdf', echo=TRUE, warning=FALSE, message=FALSE, error=TRUE)
```

---
title: 'Termination phenotype vs decay'
author: "Manfred Schmid"
output: html_document
---

`r format(Sys.time(), "%d %B, %Y")`


```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(RMetaTools))
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(broom))
suppressWarnings(library(knitr))
```


## Deeptools

Using deeptools to prepare heatmaps and metagene value matrices


#### deeptools matrix termination region ORF-Ts ...
```{bash, eval=FALSE}
#!/usr/bin/env bash

### ORFT TSS ###
ORFTs="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_ORFTs.bed"


cd ~/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bw/

## all bigwigs
plus_bw="*_plus*.bw"


# xPAP
# xORF-T TES pm 1k

## Mex

### xPap ip 0 rel 15
title="Mex_ip_xPappAminus"
samples=$(ls $plus_bw | grep Mex67AA_ip | grep -v 70 | grep xPappAminus)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
1000 1000 TES TESpm1k Mex67AA_ip_0 Mex67AA_ip_15 "deeptools_out/ORFT/${title}_ORFT_TESpm1k" "--binSize=50 --missingDataAsZero --averageTypeBins sum"


## Nab

### xPap ip 0 rel 15
title="Nab_ip_xPappAminus"
samples=$(ls $plus_bw | grep Nab2AA_ip | grep -v 70 | grep xPappAminus)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
1000 1000 TES TESpm1k Nab2AA_ip_0 Nab2AA_ip_15 "deeptools_out/ORFT/${title}_ORFT_TESpm1k" "--binSize=50 --missingDataAsZero --averageTypeBins sum"

```

#### deeptools matrix pA-site proximal signal of tandem genes...
```{bash, eval=FALSE}
# noPAP
# xORF-T TES pm 1k

## Mex

### noPap ip 0 rel 15
title="Mex_ip_noPap"
samples=$(ls $plus_bw | grep Mex67AA_ip | grep -v 70 | grep noPap)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
1000 1000 TES TESpm1k Mex67AA_ip_0 Mex67AA_ip_15 "deeptools_out/ORFT/${title}_ORFT_TESpm1k" "--binSize=50 --missingDataAsZero --averageTypeBins sum"


### noPap in 0 rel 15
title="Mex_input_noPap"
samples=$(ls $plus_bw | grep Mex67AA_input | grep -v 70 | grep noPap)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
1000 1000 TES TESpm1k Mex67AA_input_0 Mex67AA_input_15 "deeptools_out/ORFT/${title}_ORFT_TESpm1k" "--binSize=50 --missingDataAsZero --averageTypeBins sum"

## Nab

### noPap ip 0 rel 15
title="Nab_ip_noPap"
samples=$(ls $plus_bw | grep Nab2AA_ip | grep -v 70 | grep noPap)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
1000 1000 TES TESpm1k Nab2AA_ip_0 Nab2AA_ip_15 "deeptools_out/ORFT/${title}_ORFT_TESpm1k" "--binSize=50 --missingDataAsZero --averageTypeBins sum"


### noPap in 0 rel 15
title="Nab2_input_noPap"
samples=$(ls $plus_bw | grep Nab2AA_input | grep -v 70 | grep noPap)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
1000 1000 TES TESpm1k Nab2AA_input_0 Nab2AA_input_15 "deeptools_out/ORFT/${title}_ORFT_TESpm1k" "--binSize=50 --missingDataAsZero --averageTypeBins sum"

```


#### deeptools termination defect

this is based on Mex specifically
```{bash, eval = FALSE}
## get the termination defect ... 250bp window
title="Mex_ip_xPappAminus"
samples=$(ls $plus_bw | grep Mex67AA_ip | grep -v 70 | grep xPappAminus)
sample_names=$(echo $samples | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

bash ~/ms_tools/MS_Metagene_Tools/RNAseq_sensitivityX.sh reference-point \
$ORFTs "$samples" "${samples//plus/minus}" \
0 250 TES term250 Mex67AA_ip_0 Mex67AA_ip_15 "deeptools_out/ORFT/${title}_ORFT_TESp250" "--binSize=50 --missingDataAsZero --averageTypeBins sum"

python /Users/schmidm/ms_tools/MS_Metagene_Tools/log2_matrix.py deeptools_out/ORFT/${title}_ORFT_TESp250_joined_sensitivity.gz
 
plotHeatmap -m deeptools_out/ORFT/${title}_ORFT_TESp250_joined_sensitivity_log2.gz \
      --sortUsing max --colorMap Blues --missingDataColor white \
      --refPointLabel TES \
      -out deeptools_out/ORFT/${title}_ORFT_TESp250_sensitivity_log2_heatmap.pdf \
      --outFileSortedRegions deeptools_out/ORFT/sorted_by_Mex_term_defect_250.bed

```


## plot Heatmaps

#### log2 values
```{bash, eval = FALSE}
for mat in deeptools_out/ORFT/*_ORFT_TESpm1k_joined.gz
  do
    python /Users/schmidm/ms_tools/MS_Metagene_Tools/log2_matrix.py $mat
  done

for mat in deeptools_out/ORFT/*_ORFT_TESpm1k_joined_sensitivity.gz
  do
    python /Users/schmidm/ms_tools/MS_Metagene_Tools/log2_matrix.py $mat
  done
```


#### sort matrices by term defect
```{bash, eval = FALSE}
for mat in deeptools_out/ORFT/*_ORFT_TESpm1k_joined*log2.gz
  do
    echo $mat
    python /Users/schmidm/ms_tools/MS_Metagene_Tools/sort_matrix_using_bed.py $mat deeptools_out/ORFT/sorted_by_Mex_term_defect_250.bed -o ${mat/.gz/_sorted_by_Mex_term250.gz}
  done
```


#### plot sensitivity heatmaps
```{bash, eval = FALSE}
#plot log2 sensitivities pAplus
for mat in deeptools_out/ORFT/*_noPap_ORFT_TESpm1k_joined_sensitivity_log2_sorted_by_Mex_term250.gz
  do
    echo $mat
    plotHeatmap -m ${mat} \
      --sortRegions no --missingDataColor white \
      --refPointLabel TES \
      --zMin -3 --zMax 3 \
      --yMin -5 --yMax 5 \
      -out ${mat/.gz/_heatmap.pdf}
  done

#plot log2 sensitivities pAminus
for mat in deeptools_out/ORFT/*_xPappAminus_ORFT_TESpm1k_joined_sensitivity_log2_sorted_by_Mex_term250.gz
  do
    echo $mat
    plotHeatmap -m ${mat} \
      --sortRegions no --missingDataColor white \
      --refPointLabel TES \
      --zMin -10 --zMax 10 \
      --yMin -5 --yMax 5 \
      -out ${mat/.gz/_heatmap.pdf}
  done

```


#### decay vs termination correlation

## ie termination region sensitivity  
```{r}
path <- '/Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bw/deeptools_out/ORFT/'
file <- 'Mex_ip_xPappAminus_ORFT_TESpm1k_joined_sensitivity_log2.gz'

(mex_termination <- RMetaTools::load_deeptoolsmatrix(paste0(path,file)) %>%
    filter(rel_pos >= 0, rel_pos < 250) %>%
    dplyr::mutate(region = 'term_xPap') %>%
    group_by(sample_name, id, region) %>%
    summarize(log_mean_value=mean(value)))
```

```{r}
file <- 'Nab_ip_xPappAminus_ORFT_TESpm1k_joined_sensitivity_log2.gz'
(nab_termination <- RMetaTools::load_deeptoolsmatrix(paste0(path,file)) %>%
    filter(rel_pos >= 0, rel_pos < 250) %>%
    dplyr::mutate(region = 'term_xPap') %>%
    group_by(sample_name, id, region) %>%
    summarize(log_mean_value=mean(value)))
```

```{r}
ggplot(mex_termination, aes(x=log_mean_value, color=sample_name)) + geom_density()
```



#### pA+ defect  
```{r}
file <- 'Mex_ip_noPap_ORFT_TESpm1k_joined_sensitivity_log2.gz'
(mex_pas <- RMetaTools::load_deeptoolsmatrix(paste0(path,file))  %>%
    filter(rel_pos >= -50, rel_pos < 50) %>%
    dplyr::mutate(region = 'PASpm50_noPap') %>%
    group_by(sample_name, id, region) %>%
    summarize(log_mean_value=mean(value)))
```

```{r}
file <- 'Nab_ip_noPap_ORFT_TESpm1k_joined_sensitivity_log2.gz'
(nab_pas <- RMetaTools::load_deeptoolsmatrix(paste0(path,file))  %>%
    filter(rel_pos >= -50, rel_pos < 50) %>%
    dplyr::mutate(region = 'PASpm50_noPap') %>%
    group_by(sample_name, id, region) %>%
    summarize(log_mean_value=mean(value)))
```

combine  
```{r}
(all_tbl <- bind_rows(mex_termination, mex_pas) %>%
   bind_rows(., nab_termination) %>%
   bind_rows(., nab_pas) %>%
    separate(sample_name, c('strain', 'fraction', 'rapa'), sep='_') %>%
    dplyr::select(id, strain, fraction, rapa, log_mean_value, region) %>%
   spread(region, log_mean_value))
```


#### correlation
```{r Mex and Nab termination vs pA correlation}
ggplot(all_tbl, aes(x=PASpm50_noPap, y=term_xPap)) +
  geom_point(alpha=.4, size=.5) +
  facet_grid(.~strain+fraction+rapa) +
  stat_smooth(method='lm') +
  theme_bw()
```


```{r}
all_tbl %>%
  group_by(strain, fraction, rapa) %>%
  do(tidy(cor.test(.$PASpm50_noPap, .$term_xPap, method='pearson')))
```

```{r}
all_tbl %>%
  group_by(strain, fraction, rapa) %>%
  do(tidy(cor.test(.$PASpm50_noPap, .$term_xPap, method='spearman')))
```

```{r}
sessionInfo()
```
