---
title: "Correlation 10min pA+ data with decay rates and txn measures"
author: "Manfred Schmid"
output:
  pdf_document: default
  html_document: default
---

`r format(Sys.time(), "%d %B, %Y")`

```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_correlation_with_halflife_txn/'), 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```



```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(DESeq2))
```


## log2FC and DR estimates from 10min data


#### Load decay rate estimates from 10min data
```{r}
load('../../data/10min_DR_with_confint.RData', verbose=TRUE)

DR_cis
```


#### Load rapamycin-dependent log2FC from 10min data

```{r}
load('../../data/DESeq2_10min_results_df_v2.RData', verbose=TRUE)

dds_results_df
```



## Loading and comparing to published decay rate datasets

from Markdown *Published_decay_measures.Rmd*
```{r}
load(file='../../data/published_decay_estimates.RData', verbose=TRUE)

DR_literature
```


## Correlation with Decay Rates


#### combine mRNA data for the different datasets
```{r}
(LFC_spread <- dds_results_df %>%
  filter(type == 'ORF-T') %>%
  dplyr::select(id, name, type, log2FoldChange, comparison) %>%
  spread(comparison, log2FoldChange))
```

```{r}
(DR_spread <- DR_cis %>%
  ungroup %>%
  filter(rapa == 0) %>% 
  mutate(strain=paste0('DR_', strain)) %>%
  dplyr::select(strain,id,DR) %>% 
  spread(strain,DR))
```


```{r}
(mRNAs_spread <- left_join(LFC_spread, DR_spread) %>%
  left_join(., DR_literature)) 
```



#### correlation to distance fun

```{r}
cor_dist <- function(cor) {
  as.dist((1 - cor)/2)
}
```

#### spearman correlation matrix

```{r}
(spearman_cor_mat <- mRNAs_spread %>%
  dplyr::select(-name, -id, -type, -Presnyak_pA_DR) %>%
  as.matrix %>%
  na.omit %>%
  cor(., method='spearman'))
```


```{r spearman correlation txn hclust all mRNAs}
spearman_dists <- cor_dist(spearman_cor_mat)

hc <- hclust(spearman_dists)

plot(hc)
```


```{r}
spearman_cor_df <- spearman_cor_mat %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df
```

```{r}
study_order <- c("Mex_ip_15rel0", "Nab_ip_15rel0", "Mex_in_15rel0", "Nab_in_15rel0", 
              "Presnyak_total_DR", "Miller_DR", "Sun_DR",
              "DR_Mex", "DR_Nab")
```

```{r spearman correlation matrix all mRNAs}
spearman_cor_mat %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df %>%
  mutate(study1 = factor(study1, levels=study_order),
         study2 = factor(study2, levels=study_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```




### only genes sig over background

```{r}
load('../../data/DESeq2_10min_IP_rel_neg_results_df_v2.RData', verbose=TRUE)

dds_rel_neg_results_df
```


```{r}
(ip_sig_over_bkgd <- filter(dds_rel_neg_results_df,
       log2FoldChange > 0, padj < .1,
       type == 'ORF-T'
       ) %>%
  dplyr::select(name, comparison))
```


#### IPs only gene sig over background at 0

```{r}
ids_mex_sig0 <- filter(ip_sig_over_bkgd, comparison == 'Mex_ip_0relneg') %$% name

ids_nab_sig0 <- filter(ip_sig_over_bkgd, comparison == 'Nab_ip_0relneg') %$% name
  
ids_both_sig0 <- intersect(ids_mex_sig0, ids_nab_sig0)

length(ids_both_sig0)
```


```{r}
spearman_cor_mat_sig0 <- mRNAs_spread %>%
   filter(name %in% ids_both_sig0) %>%
  dplyr::select(-name, -id, -type, -Presnyak_pA_DR) %>%
  as.matrix %>%
  na.omit %>%
  cor(., method='spearman') 
```

```{r spearman correlation hclust sig over BG}
spearman_dists_sig0 <- cor_dist(spearman_cor_mat_sig0)

hc <- hclust(spearman_dists_sig0)

plot(hc)
```


```{r spearman correlation matrix only genes sig over bkgd 0}
spearman_cor_mat_sig0 %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df %>%
  mutate(study1 = factor(study1, levels=study_order),
         study2 = factor(study2, levels=study_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```



## Correlation with Transcription

load the transcription estimates from Markdown Published_txn_estimates_gene_body_counting.Rmd

```{r}
load(file='../../data/published_gene_body_txn_estimates.RData', verbose=TRUE)

txn_estimates
```


```{r}
(txn_estimates %<>%
  separate(name, c('id', 'type', 'name', 'common_name'), sep=':') %>%
  filter(type == 'ORF-T'))
```


```{r}
mRNAs_spread %<>%
   dplyr::select(-contains('DR')) %>%
   left_join(txn_estimates, .)
```


```{r}
(spearman_cor_mat <- mRNAs_spread %>%
  dplyr::select(-name, -id, -type, -common_name, -size) %>%
  as.matrix %>%
  na.omit %>%
  cor(., method='spearman'))
```


```{r spearman correlation hclust all mRNAs}
spearman_dists <- cor_dist(spearman_cor_mat)

hc <- hclust(spearman_dists)

plot(hc)
```


```{r}
spearman_cor_df <- spearman_cor_mat %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df
```

```{r}
study_order <- c("Mex_ip_15rel0", "Nab_ip_15rel0", "Mex_in_15rel0", "Nab_in_15rel0", 
                 "NETseq_mean", "CRAC_mean", "Chip2_mean", "Chipseq_mean")
```

```{r spearman correlation with txn matrix all mRNAs}
spearman_cor_mat %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df %>%
  mutate(study1 = factor(study1, levels=study_order),
         study2 = factor(study2, levels=study_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```




### only genes sig over background

```{r}
spearman_cor_mat_sig0 <- mRNAs_spread %>%
  filter(name %in% ids_both_sig0) %>%
  dplyr::select(-name, -id, -type, -common_name, -size) %>%
  as.matrix %>%
  na.omit %>%
  cor(., method='spearman') 
```

```{r spearman correlation txn hclust sig over BG}
spearman_dists_sig0 <- cor_dist(spearman_cor_mat_sig0)

hc <- hclust(spearman_dists_sig0)

plot(hc)
```


```{r spearman correlation txn matrix only genes sig over bkgd 0}
spearman_cor_mat_sig0 %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df %>%
  mutate(study1 = factor(study1, levels=study_order),
         study2 = factor(study2, levels=study_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```


```{r}
sessionInfo()
```

