---
title: "DESeq"
author: "Manfred Schmid"
output:
  pdf_document: default
  html_document: default
---
`r format(Sys.time(), "%d %B, %Y")`

```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_DESeq2_10min_data/'), 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```


```{r load packages}
library(tidyverse)
library(magrittr)
library(DESeq2)
```


## Load Data

#### for S.pombe sizeFactors

-> from file Sp_normalization.Rmd

```{r}
(Sp_10min_genes_sf <- read.table(file='../../data/Sp_10min_genes_sf.txt', 
                                 col.names=c('sample', 'Sp_sf')))
```

ups: remove the 'Pap' (ie EPAP treated test run) samples for this
```{r}
Sp_10min_genes_sf %<>% filter(!grepl('Pap', sample))
```

```{r}
Sp_sf_10min <- Sp_10min_genes_sf$Sp_sf

names(Sp_sf_10min) <- Sp_10min_genes_sf$sample

Sp_sf_10min
```



#### load and combine S. cerevisiae

ups: remove the 'Pap' (ie EPAP treated test run) samples for this


#### get and load count files
```{r load and combine S cerevisiae}
path <- '/Users/schmidm/Documents/Results/Lexogen_RNAseq/STAR_bw/KevinRoyAfiltered_bedgraph/bedtools_counts/'

sfx <- '_endpm200_count_sum_per_anno.txt$'

(file_list <- dir(path) %>% 
  keep(grepl(sfx, .) & !grepl('Pap', .)))
```

```{r}
(counts_df <-
    lapply(file_list, function(fname) {
      read.table(paste0(path, fname, sep = ''),
                 col.names =
                   c('chr', 'start', 'end', 'anno', 'sum'),
                 stringsAsFactors = FALSE) %>%
        mutate(condition = sub(sfx, '', fname) %>%
                 sub('_XUTs', '', .),
               sum = ifelse(sum == '.', 0, as.numeric(sum))) %>%
        tbl_df
    }) %>%
    bind_rows())
```


## DESeq2
for log2FC between rapamycin times I chose to do inputs and ips separately. ie the read numbers are expected to differ significantly, so dispersion estimation will be flawed, imho.


#### create all info for the DESeq2 run
```{r}
read_df <- counts_df %>%
  dplyr::select(anno, sum, condition) %>%
  spread(condition, sum) %>%
  tibble::remove_rownames(.) %>%
  data.frame %>%
  tibble::column_to_rownames(var='anno')

head(read_df)
```

```{r}
(coldata <- data.frame(sample = colnames(read_df) ) %>%
  tidyr::separate(sample, c('strain', 'fraction', 'rapa', 'rep'), by='_', remove=FALSE, fill='right') %>%
  mutate(condition = paste0(strain, '_', fraction, '_', rapa)) %>%
  tibble::remove_rownames(.) %>%
  tibble::column_to_rownames(var='sample') )
```


#### create the DESeq2 objects for input and ip
```{r}
in_ddsFullCountTable <- DESeqDataSetFromMatrix(
countData = read_df[,grepl('_in_', colnames(read_df))],
colData = coldata[coldata$fraction == 'in',],
design = ~ condition)

ip_ddsFullCountTable <- DESeqDataSetFromMatrix(
countData = read_df[,grepl('_ip_', colnames(read_df))],
colData = coldata[coldata$fraction == 'ip',],
design = ~ condition)
```


#### apply S.pombe size factors

```{r}
sizeFactors(in_ddsFullCountTable) <- Sp_sf_10min[colnames(in_ddsFullCountTable)]

sizeFactors(ip_ddsFullCountTable) <- Sp_sf_10min[colnames(ip_ddsFullCountTable)]
```

```{r}
dds_in <-  estimateDispersions(in_ddsFullCountTable)
dds_ip <-  estimateDispersions(ip_ddsFullCountTable)

dds_in <-  nbinomWaldTest(dds_in)
dds_ip <-  nbinomWaldTest(dds_ip)
```


```{r}
save(dds_in, file='../../data/DESeq2_dds_in_10min_Sp_sf.RData')
save(dds_ip, file='../../data/DESeq2_dds_ip_10min_Sp_sf.RData')
```



#### alt start point:
```{r, eval = FALSE}
load('../../data/DESeq2_dds_in_10min_Sp_sf.RData', verbose=TRUE)
load('../../data/DESeq2_dds_ip_10min_Sp_sf.RData', verbose=TRUE)
```



## Differential expression


Do the DESeq2 differential expression

#### DESeq2 results input
```{r}
comparisons <- list(
  c('Mex_in_15', 'Mex_in_0'), c('Nab_in_15', 'Nab_in_0')
)

names(comparisons) <- c('Mex_in_15rel0', 'Nab_in_15rel0')

(dds_in_results <- lapply(comparisons, function(comp) 
  results(dds_in, 
          contrast=c('condition', comp))))
```


#### DESeq2 results ip
```{r}
comparisons <- list(
  c('Mex_ip_15', 'Mex_ip_0'), c('Nab_ip_15', 'Nab_ip_0')
)

names(comparisons) <- c('Mex_ip_15rel0', 'Nab_ip_15rel0')

(dds_ip_results <- lapply(comparisons, function(comp) 
  results(dds_ip, 
          contrast=c('condition', comp))))
```

combine input and ips
```{r}
dds_results <- c(dds_in_results, dds_ip_results)
```



#### MA plots
```{r MA plots}
lapply(seq_along(dds_results), function(i) 
  plotMA(dds_results[[i]], main=names(dds_results)[i]))
```



#### Results to one tbl_df

Collect into one large dataframe
```{r}
(dds_results_df <- lapply(seq_along(dds_results), 
                          function(i) dds_results[[i]] %>% 
  data.frame %>% 
  tibble::rownames_to_column(var='anno') %>% 
  mutate(comparison = names(dds_results)[i])) %>%
  bind_rows %>%
  tbl_df)
```

separate full annotations
and fix the XUTs that were annotated differently ...
```{r}
dds_results_df %<>%
  separate(anno, c('id', 'type', 'name', 'common_name'), sep=':') %>%
  mutate(name = ifelse(id == 'XUT', type, name),
         type = ifelse(id == 'XUT', id, type),
         id = ifelse(type == 'XUT', name, id))
```


```{r}
save(dds_results_df, file='../../data/DESeq2_10min_results_df_v2.RData')
```



#### MA plots with type etc....

```{r MA plots per type}
dds_results_df %>%
  mutate(sig = (padj < .1 & !is.na(padj))) %>%
  ggplot(., aes(x=baseMean, y=log2FoldChange, color=sig)) +
  geom_point(size=.1, alpha=.5) +
  scale_x_log10() +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = c("darkgray", "red")) + 
  facet_grid(type~comparison) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=45, hjust=1),
        strip.text.x = element_text(size=6)
        )
```



## Differential enrichment IP sig over neg IP

#### comparisons
```{r}
comparisons <- list(
  c('Mex_ip_0', 'Mex_ip_neg'),
  c('Mex_ip_15', 'Mex_ip_neg'),
  c('Nab_ip_0', 'Nab_ip_neg'),
  c('Nab_ip_15', 'Nab_ip_neg')
)

names(comparisons) <- c('Mex_ip_0relneg',
  'Mex_ip_15relneg', 
  'Nab_ip_0relneg', 
  'Nab_ip_15relneg')

dds_rel_neg <- lapply(comparisons, function(comp) results(dds_ip, contrast=c('condition', comp)))
```



#### Results to one tbl_df

Collect into one large dataframe
```{r}
(dds_rel_neg_results_df <- lapply(seq_along(dds_rel_neg), function(i) dds_rel_neg[[i]] %>% 
  data.frame %>% 
  tibble::rownames_to_column(var='anno') %>% 
  mutate(comparison = names(dds_rel_neg)[i])) %>%
  bind_rows %>%
  tbl_df)
```

separate full annotations
```{r}
dds_rel_neg_results_df %<>%
  separate(anno, c('id', 'type', 'name', 'common_name'), sep=':') %>%
  mutate(name = ifelse(id == 'XUT', type, name),
         type = ifelse(id == 'XUT', id, type),
         id = ifelse(type == 'XUT', name, id))
```


```{r}
save(dds_rel_neg_results_df, file='../../data/DESeq2_10min_IP_rel_neg_results_df_v2.RData')
```



#### MA plots IP rel neg with type etc....

```{r MA plots IP rel neg per type}
dds_rel_neg_results_df %>%
  mutate(sig = (padj < .1 & !is.na(padj))) %>%
  ggplot(., aes(x=baseMean, y=log2FoldChange, color=sig)) +
  geom_point(size=.1, alpha=.5) +
  scale_x_log10() +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = c("darkgray", "red")) + 
  facet_grid(type~comparison) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle=45, hjust=1),
        strip.text.x = element_text(size=6)
        )
```

--> very good enrichment of essentially all classes and experiments...



```{r}
sessionInfo()
```

