---
title: "Mex vs Nab scatters"
author: "Manfred Schmid"
output: 
  pdf_document:
    toc: true 
    toc_depth: 3
    fig_caption: true
---

`r format(Sys.time(), "%d %B, %Y; %R")`

## Setup
```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_Mex_vs_Nab_scatters_10min_data/'), 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```


```{r load packages}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
suppressWarnings(library('knitr'))
suppressWarnings(library('broom'))
```



## load the DESeq2 results
```{r}
load(file='../../data/DESeq2_10min_results_df_v2.RData', verbose=TRUE)

dds_results_df
```

#### fix annotation types

We only consider ORF-Ts (and call them mRNA), SUTs and XUTs

```{r}
dds_results_df %<>%
  filter(type == 'ORF-T' | type == 'SUTs' | type == 'XUT' ) %>%
  mutate(type = ifelse(type == 'ORF-T', 'mRNAs', type),
         type = ifelse(type == 'XUT', 'XUTs', type),
         type = factor(type,
                       levels=c('mRNAs', 'SUTs', 'XUTs')))
```


## Mex vs Nab2 effect

```{r}
(LFC_spread <- dds_results_df %>%
  dplyr::select(id, type, log2FoldChange, comparison) %>%
  spread(comparison, log2FoldChange))
```

```{r}
scatter_theme <- list(geom_point(size=.2, alpha=.2), 
  stat_smooth(method='lm'), 
  facet_grid(.~type), 
  xlim(-4,4), ylim(-4,4),
  geom_hline(yintercept = 0, color='orange'), 
  geom_vline(xintercept = 0, color='orange'), 
  theme_bw())
```


note: no infinites in the log2FCs
```{r}
LFC_spread %>%
  filter(is.infinite(Mex_in_15rel0) | 
           is.infinite(Nab_in_15rel0) | 
           is.infinite(Mex_ip_15rel0) | 
           is.infinite(Nab_ip_15rel0))
```

but there are *NAs* of course
```{r}
LFC_spread %>%
  gather(condition, log2FC, -id, -type) %>%
  filter(is.na(log2FC)) %>%
  group_by(condition, type) %>%
  summarize(na_cnt = n())
```

but actually very few.


### inputs
```{r scatter Mex vs Nab input 15}
ggplot(LFC_spread, aes(x=Mex_in_15rel0, y=Nab_in_15rel0)) +
  scatter_theme
```

#### number of genes in input plot
```{r}
LFC_spread %>%
  filter(!is.na(Mex_in_15rel0), 
         !is.na(Nab_in_15rel0)) %>%
  group_by(type) %>%
  summarise(n = n())
```

#### mean and medians input of data shown in plot
```{r}
LFC_spread %>%
  filter(!is.na(Mex_in_15rel0), 
         !is.na(Nab_in_15rel0)) %>%
  group_by(type) %>%
  summarise(mean_mex_in = mean(Mex_in_15rel0),
            median_mex_in = median(Mex_in_15rel0),
            mean_nab_in = mean(Nab_in_15rel0), 
            median_nab_in = median(Nab_in_15rel0),
            n = n()) %>%
  kable
```


#### pearson correlation
```{r}
LFC_spread %>%
  group_by(type) %>%
  do(tidy(cor.test(.$Mex_in_15rel0, .$Nab_in_15rel0, method='pearson'))) %>%
  kable
```

#### spearman correlation
```{r}
LFC_spread %>%
  group_by(type) %>%
  do(tidy(cor.test(.$Mex_in_15rel0, .$Nab_in_15rel0, method='spearman'))) %>%
  kable
```

#### histograms
```{r}
histogram_theme <- list(geom_histogram(),
  facet_wrap(~type, scales='free'),
  xlim(-4,4),
  geom_hline(yintercept = 0, color='orange'),
  geom_vline(xintercept = 0, color='orange'),
  theme_bw())
```

```{r histograms Mex input 15}
ggplot(LFC_spread, aes(x=Mex_in_15rel0)) +
  histogram_theme
```

```{r histograms Nab input 15}
ggplot(LFC_spread, aes(x=Nab_in_15rel0)) +
  histogram_theme +
  coord_flip()
```


### IPs
```{r scatter Mex vs Nab ip 15}
ggplot(LFC_spread, aes(x=Mex_ip_15rel0, y=Nab_ip_15rel0)) +
  scatter_theme
```

#### mean and medians
```{r}
LFC_spread %>%
  filter(!is.na(Mex_ip_15rel0), !is.na(Nab_ip_15rel0)) %>%
  group_by(type) %>%
  summarise(mean_mex = mean(Mex_ip_15rel0),
            median_mex = median(Mex_ip_15rel0),
            mean_nab = mean(Nab_ip_15rel0), 
            median_nab = median(Nab_ip_15rel0),
            n = n()) %>%
  kable
```

#### pearson correlation
```{r}
LFC_spread %>%
  group_by(type) %>%
  do(tidy(cor.test(.$Mex_ip_15rel0, .$Nab_ip_15rel0, method='pearson'))) %>%
  kable
```

#### spearman correlation
```{r}
LFC_spread %>%
  group_by(type) %>%
  do(tidy(cor.test(.$Mex_ip_15rel0, .$Nab_ip_15rel0, method='spearman'))) %>%
  kable
```

#### histograms

```{r histograms Mex ip 15}
ggplot(LFC_spread, aes(x=Mex_ip_15rel0)) +
  histogram_theme
```

```{r histograms Nab ip 15}
ggplot(LFC_spread, aes(x=Nab_ip_15rel0)) +
  histogram_theme +
  coord_flip()
```


### IPs only gene sig over background


#### load sig. enrichments
```{r}
load('../../data/DESeq2_10min_IP_rel_neg_results_df_v2.RData', verbose=TRUE)

dds_rel_neg_results_df
```

clean up types as above
```{r}
dds_rel_neg_results_df %<>%
  filter(type %in% c('ORF-T', 'SUTs', 'XUT')) %>%
  mutate(type = ifelse(type == 'ORF-T', 'mRNAs', type),
         type = ifelse(type == 'XUT', 'XUTs', type),
         type = factor(type,
                       levels=c('mRNAs', 'SUTs', 'XUTs')))
```


```{r}
(ip_sig_over_bkgd <- filter(dds_rel_neg_results_df,
       log2FoldChange > 0, padj < .1,
       ) %>%
  dplyr::select(id, comparison))
```


```{r}
table(dds_rel_neg_results_df$comparison)
```

```{r}
table(ip_sig_over_bkgd$comparison)
```


#### IPs only gene sig over background at 0

```{r}
ids_mex_sig0 <- filter(ip_sig_over_bkgd, comparison == 'Mex_ip_0relneg') %$% id

ids_nab_sig0 <- filter(ip_sig_over_bkgd, comparison == 'Nab_ip_0relneg') %$% id
  
ids_both_sig0 <- intersect(ids_mex_sig0, ids_nab_sig0)

length(ids_both_sig0)
```

```{r}
LFC_spread_sig0 <- LFC_spread %>%
  filter(id %in% ids_both_sig0)
```

```{r}
table(LFC_spread$type)
```

```{r}
table(LFC_spread_sig0$type)
```


```{r scatter Mex vs Nab ip 15 sig over bkgd 0}
ggplot(LFC_spread_sig0, aes(x=Mex_ip_15rel0, y=Nab_ip_15rel0)) +
    scatter_theme
```

#### number of genes in plot
```{r}
LFC_spread_sig0 %>%
  filter(!is.na(Mex_ip_15rel0), !is.na(Nab_ip_15rel0)) %>%
  group_by(type) %>%
  summarise(n = n())
```


#### mean and medians
```{r}
LFC_spread_sig0 %>%
  filter(!is.na(Mex_ip_15rel0), !is.na(Nab_ip_15rel0)) %>%
  group_by(type) %>%
  summarise(mean_mex = mean(Mex_ip_15rel0),
            median_mex = median(Mex_ip_15rel0),
            mean_nab = mean(Nab_ip_15rel0), 
            median_nab = median(Nab_ip_15rel0),
            n = n()) %>%
  kable
```

#### pearson correlation
```{r}
LFC_spread_sig0 %>%
  group_by(type) %>%
  do(tidy(cor.test(.$Mex_ip_15rel0, .$Nab_ip_15rel0, method='pearson'))) %>%
  kable
```

#### spearman correlation
```{r}
LFC_spread_sig0 %>%
  group_by(type) %>%
  do(tidy(cor.test(.$Mex_ip_15rel0, .$Nab_ip_15rel0, method='spearman'))) %>%
  kable
```

#### histograms
```{r histograms Mex ip 15 sig over bkgd 0}
ggplot(LFC_spread_sig0, aes(x=Mex_ip_15rel0)) +
  histogram_theme
```

```{r histograms Nab ip 15 sig over bkgd 0}
ggplot(LFC_spread_sig0, aes(x=Nab_ip_15rel0)) +
  histogram_theme +
  coord_flip()
```



```{r}
sessionInfo()
```

