---
title: "Background Subtraction"
author: "Manfred Schmid"
date: "9 August 2017"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'Figures_BGsub_tracks/', dev='pdf')
```

```{r, message=FALSE,warning=FALSE}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
```


## Background subtraction for IPs

Using S.pombe normalized values: subtract values from negative IP from all IPs at single position level
ups:
--> can result in negatives!!  

need a way to subtract values at each position ...
here is a little python script that does this:
```{python subtract_bedgraph.py, eval = FALSE}
#!/usr/bin/env python
'''

subtracts values from file2 from file1

Usage: subtract_bedgraph.py file1 file2 --chunk_size --value_col -o/--outFileName

example: python subtract_bedgraph.py ChIP.bedgraph input.bedgraph --chunk_size 1000000 --value_col 5 -o ChIP_minus_input.bedgraph


'''


__author__ = 'schmidm'


import sys
import argparse
import numpy as np


parser = argparse.ArgumentParser(usage=__doc__)

parser.add_argument('file1')
parser.add_argument('file2')
parser.add_argument('--chunk_size', default=1000000, type=int, help="size of the internal array, allows for minor optimization, probably most efficient to use highest end position in bed file (optional, default = 1000000)")
parser.add_argument('--value_col', default=4, type=int, help="UPS: IMPORTANT, which column to get values for subtraction (default = 4)")
parser.add_argument('-o', '--outFileName', type=str)

args = parser.parse_args()


def print_subtracted(chr, f1_values, f2_values):
    '''
    :param chr: name of chr for output
    :param f1_values: np.array raw values
    :param f2_values: np.array values to be subtracted
    :return: nothing

    Prints the values in collapsed form to args.outFileName
    '''
    if len(f1_values) > len(f2_values):
        f2_values = np.append(f2_values, np.zeros(len(f1_values)-len(f2_values)))
    elif len(f2_values) > len(f1_values):
        f1_values = np.append(f1_values, np.zeros(len(f2_values)-len(f1_values)))

    out_values = f1_values - f2_values

    with open(args.outFileName, 'a') as outfile:
        start = 0
        value = out_values[0]
        i = 0
        while i < len(out_values):
            while i < len(out_values) and out_values[i] == value:
                i += 1
            if value != 0:
                outfile.write(chr + '\t' + str(start) + '\t' + str(i) + '\t' + str(value) + '\n')
            if i == len(out_values):
                return
            start = i
            value = out_values[i]


def read_chr(file):
    '''
    :param file: bedgraph file open for read
    :return: iterator of tuple chr, array of values
    '''
    ar = np.zeros(args.chunk_size)
    line = file.next().split('\t')
    chr = line[0]
    while True:
        try:
            while line[0] == chr:
                if int(line[2]) >= len(ar):
                    ar = np.append(ar, np.zeros(len(ar)))
                ar[int(line[1]): int(line[2])] += float(line[3])
                line = file.next().split('\t')
            yield chr, ar
            ar = np.zeros(args.chunk_size)
            chr = line[0]
        except StopIteration:
            yield chr, ar
            return


## ups: prevent appending to a pre-exisiting file -> overwrite instead !
with open(args.outFileName, 'w') as f:
    f.seek(0)
    f.truncate()


chr2=''
with open(args.file1, 'r') as file1, open(args.file2, 'r') as file2:
    f2_iter = read_chr(file2)
    for chr1, f1_values in read_chr(file1):
        if chr2 == chr1:
            print_subtracted(chr1, f1_values, f2_values)
        else:
            try:
                chr2, f2_values = f2_iter.next()
                while chr2 < chr1:
                    print_subtracted(chr2, np.zeros(len(f2_values)), f2_values)
                    chr2, f2_values = f2_iter.next()
                if chr2 == chr1:
                    print_subtracted(chr1, f1_values, f2_values)
                else:
                    print_subtracted(chr1, f1_values, np.zeros(len(f1_values)))
            except:
                f2_values = np.zeros(args.chunk_size)
                print_subtracted(chr1, f1_values, f2_values)
```



## 10min data

#### BGSub 10min track files

Here there was a separate negative IP for the Mex67AA and Nab2AA strains, but only from the 0 time point.
Use the negative control form the same strain for BGsub.

```{bash, eval=FALSE}
#!/bin/bash

#This script substracts the values at each positions covered for the native IP (neg0) from the IP tracks


cd /Users/schmidm/Documents/Results/Lexogen_RNAseq/STAR_bw/norm_and_pA_filtered_bedgraph/

## only relevant for IPs obviously

###for Mex
negip_file="norm_Mex_ip_neg_plus_KevinRoyAfiltered.bedgraph"

for f in norm_Mex_ip*plus_KevinRoyAfiltered.bedgraph
  do
    echo $f
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${f} ${negip_file} --chunk_size 100000 --value_col 4 -o ${f/.bedgraph/_BGsub.bedgraph}
    minus_fname="${f/plus/minus}"
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${minus_fname} ${negip_file/plus/minus} --chunk_size 100000 --value_col 4 -o ${minus_fname/.bedgraph/_BGsub.bedgraph}
  done


###for Nab
negip_file="norm_Nab_ip_neg_plus_KevinRoyAfiltered.bedgraph"

for f in norm_Nab_ip*plus_KevinRoyAfiltered.bedgraph
  do
    echo $f
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${f} ${negip_file} --chunk_size 100000 --value_col 4 -o ${f/.bedgraph/_BGsub.bedgraph}
    minus_fname="${f/plus/minus}"
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${minus_fname} ${negip_file/plus/minus} --chunk_size 100000 --value_col 4 -o ${minus_fname/.bedgraph/_BGsub.bedgraph}
  done

```



#### get stats for the effect of BGsub

```{bash, eval = FALSE}
cd /Users/schmidm/Documents/Results/Lexogen_RNAseq/STAR_bw/norm_and_pA_filtered_bedgraph/

#### total signals and positions above 0
echo "file sum_raw_signal sum_BGsub_signal sum_raw_positions sum_BGsub_position" > BGsubs.stats

for f in *BGsub.bedgraph
    do
        echo $f
        total=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' ${f/_BGsub/})
        BGsub=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' $f)
        total_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' ${f/_BGsub/})
        BGsub_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' $f)
        echo "${f/_BGsub.bedgraph/} $total $BGsub $total_pos $BGsub_pos" >> BGsubs.stats
    done
    
```


#### load those stats to R
```{r}
(stats_10min <- read.table('/Users/schmidm/Documents/Results/Lexogen_RNAseq/STAR_bw/norm_and_pA_filtered_bedgraph/BGsubs.stats', header=TRUE) %>% tbl_df %>%
  mutate(file = sub('^norm_', '', file) %>% 
           sub('_plus_KevinRoyAfiltered', '', .) %>%
           sub('_minus_KevinRoyAfiltered', '', .)) %>%
  group_by(file) %>%
  summarize(sum_raw_signal = sum(sum_raw_signal),
            sum_BGsub_signal = sum(sum_BGsub_signal),
            sum_raw_positions = sum(sum_raw_positions),
            sum_BGsub_position = sum(sum_BGsub_position)) %>%
  separate(file, c('strain', 'fraction', 'rapa', 'rep'), sep='_', extra='drop'))
```

--> already good that background subtraction sets negative IP to exactly 0


#### santiy check BGsub = raw - BG

get background
```{r}
(BG_stats_10min <- filter(stats_10min, rapa == 'neg', fraction == 'ip') %>%
  dplyr::rename(BG_raw_signal=sum_raw_signal) %>%
  dplyr::select(strain, fraction, BG_raw_signal) )
```


```{r}
(BGsub_Sanity_10min <- left_join(filter(stats_10min, rapa != 'neg'), BG_stats_10min) %>%
  mutate(BGsub_difference = sum_BGsub_signal - (sum_raw_signal - BG_raw_signal)) %>%
  arrange(BGsub_difference) %>%
   dplyr::select(-contains('position')))
```

overall there is a minute discrepancy, ie less than 1 normalized read difference between normalized bedgraph counts and raw bedgraph minus mock counts -> possible rounding errors, did not try to hunt this down.


#### visualize the signal stats

```{r}
(stats_10min_signal <- stats_10min %>%
  dplyr::select(-sum_raw_positions, -sum_BGsub_position) %>%
  gather(signal_type, sum_signal, -strain, -fraction, -rapa, -rep))
```

```{r}
filter(stats_10min_signal, rapa == 0)
```

```{r 10min data normalized vs normalized and BGsub signal}
filter(stats_10min_signal, rapa != 70) %>%
  ggplot(., aes(x=rep, y=sum_signal, fill=signal_type)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(fraction ~ strain+rapa) +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw()
```


#### visualize the number of genomic positions covered

note: script only counts positions above background!  
```{r}
(stats_10min_pos <- stats_10min %>%
  dplyr::select(-sum_raw_signal, -sum_BGsub_signal) %>%
  gather(signal_type, positive_positions, -strain, -fraction, -rapa, -rep))
```

```{r}
filter(stats_10min_pos, rapa == 0)
```

```{r 10min data raw vs BGsub above 0 positions}
filter(stats_10min_pos, rapa == 0) %>%
  ggplot(., aes(x=rep, y=positive_positions, fill=signal_type)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(fraction ~ strain) +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw()
```



## 2min data 

#### BGSub 2min track files

Here there was only one negative IP samples, use the noPap version and xPap version of this samples for all relevant BGsub.

```{bash, eval=FALSE}
#!/bin/bash

#This script substracts the values at each positions covered for the native IP (neg0) from the IP tracks

cd /Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bedgraph/



## for noPap samples
noPap_negip_file="norm_noPap_Mex67AA_ip_neg0_1_plus_KevinRoyAfiltered.bedgraph"

for f in norm_noPap_*ip*plus_KevinRoyAfiltered.bedgraph
  do
    echo $f
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${f} ${noPap_negip_file} --chunk_size 100000 --value_col 4 -o ${f/.bedgraph/_BGsub.bedgraph}
    fminus=${f/plus/minus}
    echo $fminus
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${fminus} ${noPap_negip_file/plus/minus} --chunk_size 100000 --value_col 4 -o ${fminus/.bedgraph/_BGsub.bedgraph}
  done



## for xPap samples
xPap_negip_file="norm_xPap_Mex67AA_ip_neg0_1_plus_KevinRoyAfiltered.bedgraph"

for f in norm_xPap_*ip*plus_KevinRoyAfiltered.bedgraph
  do
    echo $f
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${f} ${xPap_negip_file} --chunk_size 100000 --value_col 4 -o ${f/.bedgraph/_BGsub.bedgraph}
    fminus=${f/plus/minus}
    echo $fminus
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${fminus} ${xPap_negip_file/plus/minus} --chunk_size 100000 --value_col 4 -o ${fminus/.bedgraph/_BGsub.bedgraph}
  done
```



### get stats for the effect of BGsub

```{bash, eval = FALSE}
cd /Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bedgraph/

#### total signals and positions above 0
echo "file sum_raw_signal sum_BGsub_signal sum_raw_positions sum_BGsub_position" > BGsubs.stats

for f in norm_noPap*BGsub.bedgraph
    do
        echo $f
        total=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' ${f/_BGsub/})
        BGsub=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' $f)
        total_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' ${f/_BGsub/})
        BGsub_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' $f)
        echo "${f/_BGsub.bedgraph/} $total $BGsub $total_pos $BGsub_pos" >> BGsubs.stats
    done
    
for f in norm_xPap*BGsub.bedgraph
    do
        echo $f
        total=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' ${f/_BGsub/})
        BGsub=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' $f)
        total_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' ${f/_BGsub/})
        BGsub_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' $f)
        echo "${f/_BGsub.bedgraph/} $total $BGsub $total_pos $BGsub_pos" >> BGsubs.stats
    done
```


#### load those stats to R

```{r}
(stats_2min <- read.table('/Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bedgraph/BGsubs.stats', header=TRUE) %>% tbl_df %>%
  mutate(file = sub('^norm_', '', file) %>% 
           sub('_plus_KevinRoyAfiltered', '', .) %>%
           sub('_minus_KevinRoyAfiltered', '', .)) %>%
  group_by(file) %>%
  summarize(sum_raw_signal = sum(sum_raw_signal),
            sum_BGsub_signal = sum(sum_BGsub_signal),
            sum_raw_positions = sum(sum_raw_positions),
            sum_BGsub_position = sum(sum_BGsub_position)) %>%
  separate(file, c('Pap', 'strain', 'fraction', 'rapa', 'rep'), sep='_', extra='drop'))
```

--> again good that background subtraction sets negative IP to exactly 0


#### santiy check BGsub = raw - BG
```{r}
(BG_stats_2min <- filter(stats_2min, rapa == 'neg0', fraction == 'ip') %>%
  dplyr::rename(BG_raw_signal=sum_raw_signal) %>%
  dplyr::select(Pap, fraction, BG_raw_signal) )
```


```{r}
(BGsub_Sanity_2min <- left_join(filter(stats_2min, rapa != 'neg0'), BG_stats_2min) %>%
  mutate(BGsub_difference = sum_BGsub_signal - (sum_raw_signal - BG_raw_signal)) %>%
  arrange(BGsub_difference) %>%
   dplyr::select(-contains('position')))
```

overall there is a minute discrepancy, ie less than 1 normalized read difference between normalized bedgraph counts and raw bedgraph minus mock counts -> possible rounding errors, did not try to hunt this down.



#### visualize the signal stats

```{r}
(stats_2min_signal <- stats_2min %>%
  dplyr::select(-sum_raw_positions, -sum_BGsub_position) %>%
  gather(signal_type, sum_signal, -Pap, -strain, -fraction, -rapa, -rep))
```

```{r}
filter(stats_2min_signal, rapa == 0)
```

```{r 2min data raw vs BGsub signal}
filter(stats_2min_signal, rapa == 0 | rapa == 'neg0') %>%
  ggplot(., aes(x=rep, y=sum_signal, fill=signal_type)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(fraction ~ Pap+strain+rapa) +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw()
```


#### visualize the number of genomic positions covered

note: script only counts positions above background!  
```{r}
(stats_2min_pos <- stats_2min %>%
  dplyr::select(-sum_raw_signal, -sum_BGsub_signal) %>%
  gather(signal_type, positive_positions, -Pap, -strain, -fraction, -rapa, -rep))
```

```{r}
filter(stats_2min_pos, rapa == 0)
```

```{r 2min data raw vs BGsub above 0 positions}
filter(stats_2min_pos, rapa == 0) %>%
  ggplot(., aes(x=rep, y=positive_positions, fill=signal_type)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(fraction ~ Pap+strain) +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw()
```




#### some extra sanity check:
```{bash, eval = FALSE}
### sanity check

awk '{sum += $4*($3-$2)}END{printf("[%8.8f]\n",sum)}' norm_noPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered.bedgraph
#[259324.38465475]
awk '{sum += $4*($3-$2)}END{printf("[%8.8f]\n",sum)}' norm_noPap_Mex67AA_ip_neg0_1_plus_KevinRoyAfiltered.bedgraph
#[180113.65800485]
awk '{sum += $4*($3-$2)}END{printf("[%8.8f]\n",sum)}' norm_noPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bedgraph
#[79210.72665005]


awk '{sum += $4*($3-$2)}END{printf("[%8.8f]\n",sum)}' norm_noPap_Mex67AA_ip_0_1_minus_KevinRoyAfiltered.bedgraph
#[249558.06504378]
awk '{sum += $4*($3-$2)}END{printf("[%8.8f]\n",sum)}' norm_noPap_Mex67AA_ip_neg0_1_minus_KevinRoyAfiltered.bedgraph
#[169797.62927088]
awk '{sum += $4*($3-$2)}END{printf("[%8.8f]\n",sum)}' norm_noPap_Mex67AA_ip_0_1_minus_KevinRoyAfiltered_BGsub.bedgraph
#[79760.43577305]



##sanity check stats around SNR4 3' end

ip_file="norm_noPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered.bedgraph"
awk '$1=="chrV" && $2 == 424882' $ip_file
#chrV	424882	424883	22.6875


noPap_negip_file="norm_noPap_Mex67AA_ip_neg0_1_plus_KevinRoyAfiltered.bedgraph"
awk '$1=="chrV" && $2 == 424882' $noPap_negip_file
#chrV	424882	424883	134.291

awk '$1=="chrV" && $2 == 424882' ${ip_file/.bedgraph/_BGsub.bedgraph}
#chrV	424882	424883	-111.6035



ip_file2="norm_xPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered.bedgraph"
awk '$1=="chrV" && $2 == 424882' $ip_file2
#chrV	424882	424883	44904.3

xPap_negip_file="norm_xPap_Mex67AA_ip_neg0_1_plus_KevinRoyAfiltered.bedgraph"
awk '$1=="chrV" && $2 == 424882' $xPap_negip_file
#chrV	424882	424883	44102.1

awk '$1=="chrV" && $2 == 424882' ${ip_file2/.bedgraph/_BGsub.bedgraph}
#chrV	424882	424883	802.2



ip_file3="norm_xPap_Nab2AA_ip_0_1_plus_KevinRoyAfiltered.bedgraph"
awk '$1=="chrV" && $2 == 424882' $ip_file3
#chrV	424882	424883	29549.6

awk '$1=="chrV" && $2 == 424882' ${ip_file3/.bedgraph/_BGsub.bedgraph}
#chrV	424882	424883	-14552.5

```

--> everything behaves as expected, seems to be OK..
--> only problem perhaps the large negative values for certain positions ...


```{r}
sessionInfo()
```



