---
title: "Published txn estimates, mean counts per gene body"
author: "Manfred Schmid"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_published_txn_estimates_gene_body_counting/'), 
                      dev='pdf', echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```

`r format(Sys.time(), "%d %B, %Y; %R")`


```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(knitr))
suppressWarnings(library(broom))
```

The idea here is that different techniques measure RNAPII, they should have same "coverage" as our technique inside gene bodies.
On a metagene level that is the same "mean" value inside gene bodies.
So count up these for the different datasets.
I use bigWigAverageOverBed from UCSC tools for this. Here column mean0 (nr. 5) gives the mean couting non-covered bases as 0.


## NETseq

#### pre-processing
```{bash, eval = FALSE}
# data are wig files from GSE25107
#
# according to GEO: using Saccharomyces Genome Database (SGD, http://www.yeastgenome.org/) on October 11, 2009
## this would be close to sacCer2 ...


## for liftOver the closest chain file I can get is:
### V62_2009_02_18_V64_2011_02_03.over.chain

cat V62_2009_02_18_V64_2011_02_03.over.chain | sed 's/_...._.._.. / /g' > V62_2009_02_18_V64_2011_02_03_plain.over.chain

#dummy command ./liftOver anno.bed V62_2009_02_18_V64_2011_02_03_plain.over.chain converted.bed conversion_failed.bed

f="GSM617027_WT_NC_minus.wig"
head  $f | \
grep -v ^track | \
awk '{if($1=="variableStep"){chr=$2;sub("chrom=","",chr)}else{print chr"\t"$1-1"\t"$1"\t"$2}}' > test.bed

python ~/ms_tools/convert_chr_names2.py test.bed -f latin -t roman0 > testR.bed

/Users/schmidm/Documents/genomewide_datasets/scripts/liftOver_and_conversionTools/liftOver testR.bed V62_2009_02_18_V64_2011_02_03_plain.over.chain test_converted.bed conversion_failed.bed

python ~/ms_tools/convert_chr_names2.py test_converted.bed -f roman0 -t latin > test.bed
/Users/schmidm/ms_tools/bedGraphToBigWig test.bed /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/sacCer3_genome_latin.info ${f/.wig/.bw}


for f in *.wig
	do
		 cat  $f | grep -v ^track | awk '{if($1=="variableStep"){chr=$2;sub("chrom=","",chr)}else{print chr"\t"$1-1"\t"$1"\t"$2}}' > test.bed
		
		python ~/ms_tools/convert_chr_names2.py test.bed -f latin -t roman0 > testR.bed
		
		/Users/schmidm/Documents/genomewide_datasets/scripts/liftOver_and_conversionTools/liftOver testR.bed V62_2009_02_18_V64_2011_02_03_plain.over.chain test_converted.bed conversion_failed.bed

		python ~/ms_tools/convert_chr_names2.py test_converted.bed -f roman0 -t latin > test.bed
		
		/Users/schmidm/ms_tools/bedGraphToBigWig test.bed /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/sacCer3_genome_latin.info ${f/.wig/_sacCer3.bw}
	done
	
	
rm test*
rm *.wig



mkdir bw
mv *.bw bw/.

cd bw
ls
#GSM617027_WT_NC_minus_sacCer3.bw        GSM617029_RCO1D_minus_sacCer3.bw        GSM617031_EAF3D_minus_sacCer3.bw        GSM617033_SET1D_minus_sacCer3.bw
#GSM617027_WT_NC_plus_sacCer3.bw         GSM617029_RCO1D_plus_sacCer3.bw         GSM617031_EAF3D_plus_sacCer3.bw         GSM617033_SET1D_plus_sacCer3.bw
#GSM617028_WT_mRNA_minus_sacCer3.bw      GSM617030_DST1D_minus_sacCer3.bw        GSM617032_SET2D_minus_sacCer3.bw
#GSM617028_WT_mRNA_plus_sacCer3.bw       GSM617030_DST1D_plus_sacCer3.bw         GSM617032_SET2D_plus_sacCer3.bw
```


#### count
```{bash, eval = FALSE}
#!/usr/bin/env bash

cd /Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Churchman_NETseq/bw

#### gene body regions
body_plus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_plus.bed"
body_minus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_minus.bed"

plus_bw="GSM617027_WT_NC_plus_sacCer3.bw"
/Users/schmidm/ms_tools/bigWigAverageOverBed $plus_bw $body_plus_bed tmp_plus.count

/Users/schmidm/ms_tools/bigWigAverageOverBed ${plus_bw/plus/minus} $body_minus_bed tmp_minus.count

cat tmp_plus.count tmp_minus.count > ../averageOverBed/${plus_bw/_plus_sacCer3.bw/_body_endsm200_cleaned.counts}

rm tmp_*us.count
```

### load into R and save
```{r, warning=FALSE, message=FALSE}
(netseq_ucsc_means <- read_tsv('/Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Churchman_NETseq/averageOverBed/GSM617027_WT_NC_body_endsm200_cleaned.counts', col_names = c('name', 'size', 'covered', 'sum', 'mean0', 'mean')))
```

```{r}
(netseq_ucsc_means %<>%
  dplyr::select(name, size, mean0) %>%
  dplyr::rename(NETseq_mean = mean0))
```

```{r}
save(netseq_ucsc_means, file='data/NETseq_body_mean.RData')
```




## CRAC Milligan

#### pre-processing
```{bash, eval = FALSE}
#!/usr/bin/env bash

# downloaded gff files from GEO GSM1706520 and GSM1706521
# info is a bit limiting but seems to be de-duplicated etc but not corrected for mutations
# positions relative sacCer3

head GSM1706520_Rpo21_nabFT.gff
#    # novoalign (2.07.00MT - Aug  5 2010 @ 18:45:42) - A short read aligner with qualities.
#    # (C) 2008 NovoCraft
#    # Licensed for evaluation, educational, and not-for-profit use only.
#    #  novoalign -c 55 -f All_Rpo21_nabFT.fasta -r Random -d /home/bifx-rta/s0898477/Project_004_CLAMP/References/Saccharomyces_cerevisiae.EF4.68.novoindex -s 1
#    # Interpreting input files as FASTA.
#    # Index Build Version: 2.7
#    # Hash length: 11
#    # Step size: 1
#    chrIV   novo2gff        novo    1305623 1305661 .       +       .       >28_12706##GCC
#    chrVIII novo2gff        novo    133073  133111  .       -       .       >23_15410##GCC


# gff is 1-based half-open, simply convert to bedgraph needs to subtract -1 from start

genome_info="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/sacCer3_genome_latin.info"

cat GSM1706520_Rpo21_nabFT.gff | \
grep -v ^# | \
awk '{if($7=="+"){print $1"\t"$4-1"\t"$5"\t1"}}' | \
sort -k1,1 -k2,2n | \
bedtools genomecov -bg -i stdin -g ${genome_info} > GSM1706520_Rpo21_nabFT_plus.bed

cat GSM1706520_Rpo21_nabFT.gff | \
grep -v ^# | \
awk '{if($7=="-"){print $1"\t"$4-1"\t"$5"\t1"}}' | \
sort -k1,1 -k2,2n | \
bedtools genomecov -bg -i stdin -g ${genome_info} > GSM1706520_Rpo21_nabFT_minus.bed

cat GSM1706521_Rpo21-2_nabFT.gff | \
grep -v ^# | \
awk '{if($7=="+"){print $1"\t"$4-1"\t"$5"\t1"}}' | \
sort -k1,1 -k2,2n | \
bedtools genomecov -bg -i stdin -g ${genome_info} > GSM1706521_Rpo21-2_nabFT_plus.bed

cat GSM1706521_Rpo21-2_nabFT.gff | \
grep -v ^# | \
awk '{if($7=="-"){print $1"\t"$4-1"\t"$5"\t1"}}' | \
sort -k1,1 -k2,2n | \
bedtools genomecov -bg -i stdin -g ${genome_info} > GSM1706521_Rpo21-2_nabFT_minus.bed



### convert to bigwigs
mkdir bw

/Users/schmidm/ms_tools/bedGraphToBigWig GSM1706520_Rpo21_nabFT_plus.bed ${genome_info} bw/GSM1706520_Rpo21_nabFT_plus.bw

/Users/schmidm/ms_tools/bedGraphToBigWig GSM1706520_Rpo21_nabFT_minus.bed ${genome_info} bw/GSM1706520_Rpo21_nabFT_minus.bw

/Users/schmidm/ms_tools/bedGraphToBigWig GSM1706521_Rpo21-2_nabFT_plus.bed ${genome_info} bw/GSM1706521_Rpo21-2_nabFT_plus.bw

/Users/schmidm/ms_tools/bedGraphToBigWig GSM1706521_Rpo21-2_nabFT_minus.bed ${genome_info} bw/GSM1706521_Rpo21-2_nabFT_minus.bw
```


#### count
```{bash, eval = FALSE}
#!/usr/bin/env bash

cd /Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Tollervey_RNAP_CRAC/bw

#### gene body regions
body_plus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_plus.bed"
body_minus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_minus.bed"

plus_bw="GSM1706520_Rpo21_nabFT_plus.bw"
/Users/schmidm/ms_tools/bigWigAverageOverBed $plus_bw $body_plus_bed tmp_plus.count

/Users/schmidm/ms_tools/bigWigAverageOverBed ${plus_bw/plus/minus} $body_minus_bed tmp_minus.count

cat tmp_plus.count tmp_minus.count > ../averageOverBed/${plus_bw/_plus.bw/_body_endsm200_cleaned.counts}

rm tmp_*us.count
```


### load into R and save
```{r, warning=FALSE, message=FALSE}
(crac_ucsc_means <- read_tsv('/Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Tollervey_RNAP_CRAC/averageOverBed/GSM1706520_Rpo21_nabFT_body_endsm200_cleaned.counts', col_names = c('name', 'size', 'covered', 'sum', 'mean0', 'mean')))
```

```{r}
(crac_ucsc_means %<>%
  dplyr::select(name, size, mean0) %>%
  dplyr::rename(CRAC_mean = mean0))
  
```

```{r}
save(crac_ucsc_means, file='data/CRAC_body_mean.RData')
```


## compare CRAC and NETseq

```{r CRAC vs NETseq scatter}
left_join(netseq_ucsc_means, crac_ucsc_means) %>%
  ggplot(., aes(x=CRAC_mean, y=NETseq_mean)) + geom_point() + scale_x_log10() + scale_y_log10()
```


correlation, only genes positive in both:  
```{r}
left_join(netseq_ucsc_means, crac_ucsc_means) %>%
  filter(NETseq_mean > 0, CRAC_mean > 0) %>%
  do(tidy(cor.test(log2(.$CRAC_mean), log2(.$NETseq_mean), use='pairwise.complete')))
```


## Cramer Rpb3 chip-on-chip

#### pre-processing
We are taking data from Cramer labs Rpb3 chip on chip from ArrayExpress E-TABM-1033.  

Raw file is a wig file, but intervals are overlapping, and its based on sacCer2 so takes a bit of pre-processing.  
```{bash, eval = FALSE}
#!/usr/bin/env bash


## file Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.wig is downloaded from ArrayExpress E-TABM-1033
## this is Steinmetz tiling array format


head Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.wig
#track type=wiggle_0 name=Rpb3_InputAndMockNormalized
#variableStep chrom=chrV span=25
#1 0.108703797559
#5 0.109159129901
#9 0.108703797559
#13 0.109159129901
#17 0.108703797559
#21 0.109159129901
#25 0.109614462242
#29 0.112107222775

##largest position on each chromosome...
awk '{
  if($1 ~/^variableStep/ ){
    print max_pos
    print $0
    max_pos=0
  }else{
    max_pos=$1
  }
}END{
  print max_pos
}' Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.wig | \
sed 1d | paste - - | \
sed s/"variableStep "//g | sed s/span=25//g | sed s/chrom=//g | sort -k1,1
#chrI    230187
#chrII   813156
#chrIII  316590
#chrIV   1531896
#chrIX   439861
#chrV    576846
#chrVI   270125
#chrVII  1090919
#chrVIII         562616
#chrX    745646
#chrXI   666426
#chrXII  1078148
#chrXIII         924405
#chrXIV  784312
#chrXV   1091263
#chrXVI  948039
#
#chr		sacCer1	sacCer2	sacCer3	##found max position(-25) on tiling wig data
#chr01	230208	230208	230218	##230187
#chr02	813136	813178	813184	##813156
#chr03	316613	316617	316620	##316590
#chr04	1531914	1531919	1531933	##1531896
#chr05	576869	576869	576874	##576846
#chr06	270148	270148	270161	##270125
#chr07	1090944	1090947	1090940	##1090919
#chr08	562639	562643	562643	##562616
#chr09	439885	439885	439888	##439861
#chr10	745446	745742	745751	##745646 -->!!
#chr11	666445	666454	666816	##666426
#chr12	1078173	1078175	1078177	##1078148
#chr13	924430	924429	924431	##924405
#chr14	784328	784333	784333	##784312
#chr15	1091285	1091289	1091291	##1091263
#chr16	948060	948062	948066	##948039
#chr17	85779	85779	85779


##genome version closely related to sacCer2
## lift to sacCer3
##UCSC sacCer2 = S288C_reference_genome_R61-1-1_20080605.tgz
##UCSC sacCer3 = S288C_reference_genome_R64-1-1_20110203.tgz


/Users/schmidm/ms_tools/wigToBigWig -clip Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.wig /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer2/sacCer2_genome_latin.info Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.bw



## for liftOver the chain file for sacCer2 to sacCer3 is :
### V61_2008_06_05_V64_2011_02_03.over.chain
cat /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer2/V61_2008_06_05_V64_2011_02_03.over.chain | sed 's/_...._.._.. / /g' > V61_2008_06_05_V64_2011_02_03_plain.over.chain

#dummy command ./liftOver anno.bed V62_2009_02_18_V64_2011_02_03_plain.over.chain converted.bed conversion_failed.bed

f="Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.wig"

cat $f | \
grep -v ^track | \
awk '{
  if($1=="variableStep"){
    if(chr!=""){
      for(i=1;i<=max_pos;i++){
        if(n[i]>0){
          print chr"\t"i-1"\t"i"\t"ar[i]/n[i]
        }
        ar[i]=0
        n[i]=0
      }
    }
    chr=$2;
    sub("chrom=","",chr);
    span=$3;
    sub("span=","",span);
  }else{
    max_pos=$1+span
    for(i=$1;i<=($1+span);i++){
      n[i]+=1
      ar[i]+=$2
    }
  }
}END{
  for(i=1;i<=max_pos;i++){
    if(n[i]>0){
      print chr"\t"i-1"\t"i"\t"ar[i]/n[i]
    }
  }
}' > ${f/.wig/.bedGraph}

python ~/ms_tools/convert_chr_names2.py ${f/.wig/.bedGraph} -f latin -t roman0 > roman.bedGraph

/Users/schmidm/Documents/genomewide_datasets/scripts/liftOver_and_conversionTools/liftOver roman.bedGraph V61_2008_06_05_V64_2011_02_03_plain.over.chain roman_converted.bed conversion_failed.bed

python ~/ms_tools/convert_chr_names2.py roman_converted.bed -f roman0 -t latin > ${f/.wig/_sacCer3.bedGraph}

sort -k1,1 -k2,2n ${f/.wig/_sacCer3.bedGraph} -o ${f/.wig/_sacCer3.bedGraph}

/Users/schmidm/ms_tools/bedGraphToBigWig ${f/.wig/_sacCer3.bedGraph} /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/sacCer3_genome_latin.info ${f/.wig/_sacCer3.bw}

rm Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized_sacCer3.bedGraph
rm Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized.bedGraph
rm roman*

```


#### count

```{bash, eval = FALSE}
#!/usr/bin/env bash

cd /Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Cramer_ChIPchip

#### gene body regions
body_plus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_plus.bed"
body_minus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_minus.bed"

bw="Cramer_E_TABM_1033_Rpb3_InputAndMockNormalized_sacCer3.bw"
/Users/schmidm/ms_tools/bigWigAverageOverBed $bw $body_plus_bed tmp_plus.count

/Users/schmidm/ms_tools/bigWigAverageOverBed $bw $body_minus_bed tmp_minus.count

cat tmp_plus.count tmp_minus.count > averageOverBed/${bw/_InputAndMockNormalized_sacCer3.bw/_body_endsm200_cleaned.counts}

rm tmp_*us.count
```


#### load into R and save
```{r, warning=FALSE, message=FALSE}
(rpb3_chip2_ucsc_means <- read_tsv('/Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Cramer_ChIPchip/averageOverBed/Cramer_E_TABM_1033_Rpb3_body_endsm200_cleaned.counts', col_names = c('name', 'size', 'covered', 'sum', 'mean0', 'mean')))
```

```{r}
(rpb3_chip2_ucsc_means %<>%
  dplyr::select(name, size, mean0) %>%
  dplyr::rename(Chip2_mean = mean0))
  
```

```{r}
save(rpb3_chip2_ucsc_means, file='data/Rpb3ChIP2_body_mean.RData')
```


#### compare Rpb3 ChIP^2 and NETseq

```{r ChIP^2 vs NETseq scatter}
left_join(netseq_ucsc_means, rpb3_chip2_ucsc_means) %>%
  ggplot(., aes(x=Chip2_mean, y=NETseq_mean)) + geom_point() + scale_x_log10() + scale_y_log10()
```



## Henikoff Rpb3-FLAG ChIPseq

We are taking data from Henikoffs lab ChIPseq for Rpb3 sample GSM2551210 from GSE97081.  

#### pre-processing
Raw file is a wig file mapped to sacCer3, though some intervals exceeding right end of each chromosome afaics. Did not investigate this further and used data as-is.  
Processing:  
```{bash, eval = FALSE}
#!/usr/bin/env bash

cd /Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Henikoff_Rpb3FLAG_ChIPseq

## -> mapped to sacCer3, mentioned specifically at GEO
## data is coverage normalized to S.pombe spike-ins

awk '{
  if($1 ~/^variableStep/ ){
    print max_pos
    print $0
    max_pos=0
  }else{
    max_pos=$1
  }
}END{
  print max_pos
}' GSM2551210_WT_A_140.wig | \
sed 1d | paste - - | \
sed s/"variableStep "//g | sed s/span=1//g | sed s/chrom=//g | sort -k1,1

# chrI   230217
# chrII  813185
# chrIII         316616
# chrIV  1531912
# chrIX  439889
# chrM   85780
# chrV   576871
# chrVI  270162
# chrVII         1090936
# chrVIII        562650
# chrX   745754
# chrXI  666817
# chrXII         1078177
# chrXIII        924431
# chrXIV         784338
# chrXV  1091298
# chrXVI         948062


## only cerevisiae chromosomes in this file
/Users/schmidm/ms_tools/wigToBigWig GSM2551210_WT_A_140.wig /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/sacCer3_genome_latin.info GSM2551210_WT_A_140.bw
## complains, look like wig file contains entries off chromosome end, just a few nts but still !!

## also chrM not found in genome
## -> fix this


cat /Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/sacCer3_genome_latin.info | sed s/Mito/chrM/g > sacCer3_genome.info


/Users/schmidm/ms_tools/wigToBigWig -clip GSM2551210_WT_A_140.wig sacCer3_genome.info GSM2551210_WT_A_140.bw
#line 1959084 of GSM2551210_WT_A_140.wig: chromosome chrIX has 439888 bases, but item ends at 439889
#line 3047743 of GSM2551210_WT_A_140.wig: chromosome chrXIV has 784333 bases, but item ends at 784334
#line 3047744 of GSM2551210_WT_A_140.wig: chromosome chrXIV has 784333 bases, but item ends at 784335
#line 3047745 of GSM2551210_WT_A_140.wig: chromosome chrXIV has 784333 bases, but item ends at 784336
#line 3047746 of GSM2551210_WT_A_140.wig: chromosome chrXIV has 784333 bases, but item ends at 784337
#line 3047747 of GSM2551210_WT_A_140.wig: chromosome chrXIV has 784333 bases, but item ends at 784338
#line 4201663 of GSM2551210_WT_A_140.wig: chromosome chrM has 85779 bases, but item ends at 85780
#line 5166715 of GSM2551210_WT_A_140.wig: chromosome chrX has 745751 bases, but item ends at 745752
#line 5166716 of GSM2551210_WT_A_140.wig: chromosome chrX has 745751 bases, but item ends at 745753
#line 5166717 of GSM2551210_WT_A_140.wig: chromosome chrX has 745751 bases, but item ends at 745754
#line 8110064 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091292
#line 8110065 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091293
#line 8110066 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091294
#line 8110067 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091295
#line 8110068 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091296
#line 8110069 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091297
#line 8110070 of GSM2551210_WT_A_140.wig: chromosome chrXV has 1091291 bases, but item ends at 1091298
#line 8669012 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562644
#line 8669013 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562645
#line 8669014 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562646
#line 8669015 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562647
#line 8669016 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562648
#line 8669017 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562649
#line 8669018 of GSM2551210_WT_A_140.wig: chromosome chrVIII has 562643 bases, but item ends at 562650
#line 8934728 of GSM2551210_WT_A_140.wig: chromosome chrVI has 270161 bases, but item ends at 270162
#line 9597240 of GSM2551210_WT_A_140.wig: chromosome chrXI has 666816 bases, but item ends at 666817
#line 10400860 of GSM2551210_WT_A_140.wig: chromosome chrII has 813184 bases, but item ends at 813185

```



#### count  
```{bash, eval = FALSE}
#!/usr/bin/env bash

cd /Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Henikoff_Rpb3FLAG_ChIPseq

#### gene body regions
body_plus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_plus.bed"
body_minus_bed="/Users/schmidm/Documents/genomewide_datasets/annotations/sacCer3/Steinmetz_sacCer3_latin_body_endsm200_cleaned_minus.bed"

bw="GSM2551210_WT_A_140.bw"
/Users/schmidm/ms_tools/bigWigAverageOverBed $bw $body_plus_bed tmp_plus.count

/Users/schmidm/ms_tools/bigWigAverageOverBed $bw $body_minus_bed tmp_minus.count

cat tmp_plus.count tmp_minus.count > averageOverBed/${bw/_A_140.bw/_body_endsm200_cleaned.counts}

rm tmp_*us.count
```


### load into R and save
```{r, warning=FALSE, message=FALSE}
(rpb3_chipseq_ucsc_means <- read_tsv('/Users/schmidm/Documents/genomewide_datasets/ChIP_CLiP_yeast/Henikoff_Rpb3FLAG_ChIPseq/averageOverBed/GSM2551210_WT_body_endsm200_cleaned.counts', col_names = c('name', 'size', 'covered', 'sum', 'mean0', 'mean')))
```

```{r}
(rpb3_chipseq_ucsc_means %<>%
  dplyr::select(name, size, mean0) %>%
  dplyr::rename(Chipseq_mean = mean0))
  
```

```{r}
save(rpb3_chipseq_ucsc_means, file='data/Rpb3ChIPseq_body_mean.RData')
```


#### compare Rpb3 ChIPseq and NETseq

```{r ChIPseq vs NETseq scatter}
left_join(netseq_ucsc_means, rpb3_chipseq_ucsc_means) %>%
  ggplot(., aes(x=Chipseq_mean, y=NETseq_mean)) + geom_point() + scale_x_log10() + scale_y_log10()
```


#### compare Rpb3 ChIPseq and ChIP^2

```{r ChIPseq vs ChIP1 scatter}
left_join(rpb3_chip2_ucsc_means, rpb3_chipseq_ucsc_means) %>%
  ggplot(., aes(x=Chipseq_mean, y=Chip2_mean)) + geom_point() + scale_x_log10() + scale_y_log10()
```


## combine all in one tbl

```{r}
(txn_estimates <- left_join(rpb3_chip2_ucsc_means, rpb3_chipseq_ucsc_means) %>%
  left_join(., netseq_ucsc_means) %>%
  left_join(., crac_ucsc_means))
```


```{r}
save(txn_estimates, file='../data/published_gene_body_txn_estimates.RData')
```

```{r}
sessionInfo()
```
