---
title: 'snRNA'
author: "Manfred Schmid"
output: pdf_document
---

`r format(Sys.time(), "%d %B, %Y")`

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_snRNA/'), 
                      dev='pdf', echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```


```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(knitr))
suppressWarnings(library(RMetaTools))
suppressWarnings(library(broom))
```


## heatmaps

```{bash, eval = FALSE}
### snRNA TES ###
sgdSNR="/Users/schmidm//Documents/genomewide_datasets/annotations/sacCer3/sgdSNRs_sacCer3.bed"

cd ~/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bw/

##rapa 0 rep 1 samples
plus_bw="*_0_1*_plus*.bw"
minus_bw=${plus_bw/plus/minus}

sample_names=$(echo $plus_bw | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)

## matrix
bash ~/ms_tools/MS_Metagene_Tools/computeMatrixStrandedX.sh reference-point \
$sgdSNR "$plus_bw" "$minus_bw" \
1000 1000 TES deeptools_out/SNR/snRNA_0_1_TESpm1k "--binSize=10 --missingDataAsZero"

#log2 instead of counts
python /Users/schmidm/ms_tools/MS_Metagene_Tools/log2_matrix.py deeptools_out/SNR/snRNA_0_1_TESpm1k_joined.gz

plotHeatmap -m deeptools_out/SNR/snRNA_0_1_TESpm1k_joined_log2.gz \
      --sortUsing sum --colorMap Blues --missingDataColor white \
      --refPointLabel TES \
      --samplesLabel $sample_names \
      -out deeptools_out/SNR/snRNA_0_1_TESpm1k_joined_log2_heatmap.pdf

```


## single nucleotide metagene analysis

```{bash, eval = FALSE}
### snRNA TES ###
sgdSNR="~/Documents/genomewide_datasets/annotations/sacCer3/sgdSNRs_sacCer3.bed"
cd ~/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bw/
##rapa 0 rep 1 samples
plus_bw="*_0_*_plus*.bw"
minus_bw=${plus_bw/plus/minus}
sample_names=$(echo $plus_bw | sed s/norm_//g | sed s/_BGsub//g | sed s/_plus_KevinRoyAfiltered.bw//g)


ls $plus_bw 
# norm_noPap_Mex67AA_input_0_1_plus_KevinRoyAfiltered.bw
# norm_noPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bw
# norm_noPap_Nab2AA_input_0_1_plus_KevinRoyAfiltered.bw
# norm_noPap_Nab2AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPap_Mex67AA_input_0_1_plus_KevinRoyAfiltered.bw
# norm_xPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPap_Mex67AA_ip_0_2_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPap_Mex67AA_ip_0_3_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPap_Nab2AA_input_0_1_plus_KevinRoyAfiltered.bw
# norm_xPap_Nab2AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPap_Nab2AA_ip_0_2_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPap_Nab2AA_ip_0_3_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPappAminus_Mex67AA_input_0_1_plus_KevinRoyAfiltered.bw
# norm_xPappAminus_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPappAminus_Mex67AA_ip_0_2_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPappAminus_Mex67AA_ip_0_3_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPappAminus_Nab2AA_input_0_1_plus_KevinRoyAfiltered.bw
# norm_xPappAminus_Nab2AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPappAminus_Nab2AA_ip_0_2_plus_KevinRoyAfiltered_BGsub.bw
# norm_xPappAminus_Nab2AA_ip_0_3_plus_KevinRoyAfiltered_BGsub.bw


echo $sample_names 
#noPap_Mex67AA_input_0_1 noPap_Mex67AA_ip_0_1 noPap_Nab2AA_input_0_1 noPap_Nab2AA_ip_0_1 xPap_Mex67AA_input_0_1 xPap_Mex67AA_ip_0_1 xPap_Mex67AA_ip_0_2 xPap_Mex67AA_ip_0_3 xPap_Nab2AA_input_0_1 xPap_Nab2AA_ip_0_1 xPap_Nab2AA_ip_0_2 xPap_Nab2AA_ip_0_3 xPappAminus_Mex67AA_input_0_1 xPappAminus_Mex67AA_ip_0_1 xPappAminus_Mex67AA_ip_0_2 xPappAminus_Mex67AA_ip_0_3 xPappAminus_Nab2AA_input_0_1 xPappAminus_Nab2AA_ip_0_1 xPappAminus_Nab2AA_ip_0_2 xPappAminus_Nab2AA_ip_0_3

## matrix
bash ~/ms_tools/MS_Metagene_Tools/computeMatrixStrandedX.sh reference-point \
$sgdSNR "$plus_bw" "$minus_bw" \
10 10 TES deeptools_out/SNR/snRNA_0_all_TESpm10bp_single_nucres "--binSize=1 --missingDataAsZero"

#log2 instead of counts
python ~/ms_tools/MS_Metagene_Tools/log2_matrix.py deeptools_out/SNR/snRNA_0_all_TESpm10bp_single_nucres_joined.gz

plotHeatmap -m deeptools_out/SNR/snRNA_0_all_TESpm10bp_single_nucres_joined_log2.gz \
      --sortUsing sum --colorMap Blues --missingDataColor white \
      --refPointLabel TES \
      --samplesLabel $sample_names \
      -out deeptools_out/SNR/snRNA_0_all_TESpm10bp_single_nucres_joined_log2_heatmap.pdf
```


#### single nucl to R
```{r}
snrna <- RMetaTools::load_deeptoolsmatrix('~/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bw/deeptools_out/SNR/snRNA_0_all_TESpm10bp_single_nucres_joined.gz')

snrna
```

add pseudount for IPs and pAminus at or below background
```{r}
(pseudocount <- min(snrna$value[snrna$value>0]))
```

```{r}
(snrna %<>%
  mutate(value = ifelse(value <= 0, pseudocount, value+pseudocount),
          rel_pos = rel_pos + 1,
         sample_name = sub('norm_', '', sample_name) %>% 
            sub('_BGsub', '', .) %>% 
            sub('_KevinRoyAfiltered', '', .)) %>%
   separate(sample_name, c('Pap', 'strain', 'fraction', 'rapa', 'rep'), sep='_') %>%
   mutate(pA = case_when(.$Pap == 'noPap' ~ 'pA+', 
                         .$Pap == 'xPap' ~ 'pA+/-',
                         .$Pap == 'xPappAminus' ~ 'pA-'),
          pA = factor(pA, levels=c('pA+', 'pA+/-', 'pA-'))))
```



#### average replicate and strains
```{r}
(snrna_avg <- snrna %>%
  group_by(id, name, rel_pos, pA, fraction, rapa) %>%
  summarize(value = mean(value)))
```


#### metagene values
```{r}
(snrna_meta <- snrna_avg %>%
  group_by(rel_pos, pA, fraction, rapa) %>%
  do(tidy(summary(.$value))))
```



#### plots
```{r snRNA single nuc mean and median line }
snrna_meta %>%
  filter(rapa == '0') %>%
ggplot(., aes(x=rel_pos, y=mean)) + 
  geom_line() +
  geom_line(aes(y=median), color='red') +
  scale_y_log10() +
  facet_grid(.~pA+fraction+rapa, scales='free') +
  theme_bw()
```

n for those:
```{r}
snrna_avg %>%
  group_by(rel_pos, pA, fraction, rapa) %>%
  filter(!is.na(value)) %>%
  summarise(cnt = n()) %>%
  spread(rel_pos, cnt)
```


```{r snRNA single nuc mean and median line and points individual libraries}
snrna %>%
  filter(rapa == '0') %>%
ggplot(., aes(x=rel_pos, y=value)) + 
  geom_point(size=.5) +
  geom_line(data=snrna_meta, aes(x=rel_pos, y=mean)) +
  geom_line(data=snrna_meta, aes(x=rel_pos, y=median), color='red') +
  scale_y_log10() +
  facet_grid(strain~pA+fraction+rapa+rep, scales='free') +
  theme_bw()
```



## Decay rates from those data

same calculations as above but consider genes at background as 0 decay rate for median calculation later on
```{r}
(DR_snRNAs <- snrna %>%
  filter(rapa == 0) %>%
  group_by(id, rel_pos, pA, fraction, rapa) %>%
  summarize(mean_C = mean(value)) %>%
  spread(fraction, mean_C) %>%
  mutate(ip_in_ratio = ip/input,
         DR = -log(1-ip_in_ratio)/2,
         DR = case_when((ip == pseudocount) & (input > pseudocount) ~ 0,
                        input == pseudocount ~ as.numeric(NA),
                        (ip > pseudocount) & (input > pseudocount) ~ DR)))
```

```{r DR violin with box all positions summed}
DR_snRNAs %>%
  ggplot(., aes(x=pA, y=DR, fill=pA)) +
  geom_violin() +
  geom_boxplot(width=.2, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  scale_fill_brewer(palette='Set1') +
  theme_bw() 
```

```{r DR violin with box pAminus per positions summed}
DR_snRNAs %>%
  filter(pA == 'pA-') %>%
  ggplot(., aes(x=as.factor(rel_pos), y=DR)) +
  geom_violin(fill='purple') +
  facet_grid(.~pA) +
  geom_boxplot(aes(x=as.factor(rel_pos), y=DR),
               width=.1, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  scale_fill_brewer(palette='Set1') +
  theme_bw() +
  facet_grid(.~pA)
```

```{r DR violin with box pAplus per positions summed}
DR_snRNAs %>%
  filter(pA == 'pA+') %>%
  ggplot(., aes(x=as.factor(rel_pos), y=DR)) +
  geom_violin(fill='purple') +
  facet_grid(.~pA) +
  geom_boxplot(aes(x=as.factor(rel_pos), y=DR),
               width=.1, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  scale_fill_brewer(palette='Set1') +
  theme_bw() +
  facet_grid(.~pA)
```


```{r DR boxplot and jitter pA plus and pAminus per position with y scale as major Figure}
DR_snRNAs %>%
  filter(pA == 'pA-' | pA == 'pA+') %>%
  ggplot(., aes(x=as.factor(rel_pos), y=DR)) +
  geom_boxplot(width=.5, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  stat_summary(fun.y=median, geom="line", aes(group=1, color='red'), size=1)  + 
  geom_jitter(size=.1) +
  scale_y_log10() +
  facet_grid(pA~.) +
  coord_cartesian(ylim=c(.00001,10)) +
  scale_fill_brewer(palette='Set1') +
  scale_color_brewer(palette='Set1') +
  theme_bw() +
  theme(panel.grid = element_blank())
```


most stable snRNAs
```{r}
DR_snRNAs %>%
  filter(pA == 'pA-', rel_pos == 0) %>%
  arrange(DR) %>%
  kable
```

DR per position medians
```{r}
DR_snRNAs %>%
  group_by(pA, rel_pos) %>%
  filter(!is.na(DR)) %>%
  summarize(median_DR = median(DR)) %>%
  spread(pA, median_DR)
```

```{r}
save(DR_snRNAs, file='../../data/2min_DR_snRNAs.RData')
```


```{r}
sessionInfo()
```
