---
title: "Count body and end signal for the various samples and tracks"
author: "Manfred Schmid"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
`r format(Sys.time(), "%d %B, %Y")`

```{r setup}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'Figures_body_end_counting/', dev='pdf')
```

```{r, message=FALSE,warning=FALSE}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
suppressWarnings(library('knitr'))
suppressWarnings(library('broom'))
```


Using scaled S.pombe normalized and BGsub data from *body_end_counting.Rmd* and *2min_body_end_counts_to_R.Rmd*.  


## Load body and end counts

these are BGsub and normalized to S.pombe.
```{r}
load('../../data/norm_2min_body_end_scaled_and_BGsub.RData', verbose = TRUE)

both_df
```

How many ids per annotation type:  
```{r}
distinct(both_df, type, id) %>% 
  group_by(type) %>% 
  summarize(cnt=n())
```



## Barplots signal at End vs body pA+ vs pA-

Global distribution of pA+ and pA- signal in gene body an gene end regions.

```{r}
(both_totals <- both_df %>%
  group_by(type, Pap, strain,fraction, rapa, part, pA, rep) %>% 
  summarize(total_Spnorm_signal = sum(sum)))
```

```{r barplot total all samples }
both_totals %>%
  group_by(part, pA, fraction, strain, rapa, rep) %>%
  summarize(all_types_total_Spnorm_signal = sum(total_Spnorm_signal)) %>%
ggplot(., aes(x=part, y=all_types_total_Spnorm_signal, fill=part)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_blank())
```

```{r barplot total all samples only ORFs}
both_totals %>%
  filter(type == 'ORF-T') %>%
  group_by(part, pA, fraction, strain, rapa, rep) %>%
  summarize(all_types_total_Spnorm_signal = sum(total_Spnorm_signal)) %>%
ggplot(., aes(x=part, y=all_types_total_Spnorm_signal, fill=part)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_blank())
```


--> ups negatives appear, possibly due to BGsub and pAminus procedures !


#### non-negative total signal

```{r}
(non_neg_totals <- both_df %>%
   mutate(sum = ifelse(sum < 0, 0, sum)) %>%
  group_by(type, Pap, strain,fraction, rapa, part, pA, rep) %>% 
  summarize(total_Spnorm_signal = sum(sum)))
```

```{r barplot non-negative totals all samples }
non_neg_totals %>%
  group_by(part, pA, fraction, strain, rapa, rep) %>%
  summarize(all_types_total_Spnorm_signal = sum(total_Spnorm_signal)) %>%
ggplot(., aes(x=part, y=all_types_total_Spnorm_signal, fill=part)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_blank())
```

```{r barplot non-negative totals all samples only ORFs}
non_neg_totals %>%
  filter(type == 'ORF-T') %>%
  group_by(part, pA, fraction, strain, rapa, rep) %>%
  summarize(all_types_total_Spnorm_signal = sum(total_Spnorm_signal)) %>%
ggplot(., aes(x=part, y=all_types_total_Spnorm_signal, fill=part)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_blank())
```


#### non-negative total signal per type

```{r barplot non-neg signal per type and part}
non_neg_totals %>%
  filter(rapa == '0', rep == '1') %>%
  ggplot(., aes(x=type, y=total_Spnorm_signal, fill=type)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~part+strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1, angle=45))
```

```{r barplot non-neg signal per type wo part}
non_neg_totals %>%
  filter(rapa == '0', rep == '1') %>%
  ggplot(., aes(x=type, y=total_Spnorm_signal, fill=type)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1, angle=45))
```

```{r barplot non-neg signal per type wo part for Figure}
non_neg_totals %>%
  filter(rapa == 0, rep == 1, 
         pA == 'pA+' | pA == 'pA-',
         type != 'other',
         type != 'XUT') %>%
  ungroup %>%
  mutate(pA = factor(pA, levels=c('pA+', 'pA+ + pA-', 'pA-', 'pA-X'))) %>%
  ggplot(., aes(x=type, y=total_Spnorm_signal, fill=type)) + 
  geom_bar(stat='identity') +
  facet_grid(pA+fraction~strain+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x=element_text(hjust=1, angle=45))
```


## pAplus vs pAminus body vs end piecharts

```{r}
(body_end_percents <- non_neg_totals %>%
   group_by(type, pA, strain, fraction, rapa, rep) %>%
   mutate(total_Spnorm_signal_percent = 100 * total_Spnorm_signal / sum(total_Spnorm_signal)))
```

```{r piechart body vs end all rapa 0 rep 1}
body_end_percents %>%
  filter(rapa == 0, rep == 1) %>%
ggplot(., aes(x=factor(1), y=total_Spnorm_signal_percent, fill=part)) + 
  geom_bar(stat='identity') +
  coord_polar(theta='y') +
  facet_grid(type+strain~pA+fraction+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_blank())
```


```{r piechart body vs end all rapa 0 rep 1 for Figure}
body_end_percents %>%
  filter(rapa == 0, rep == 1, 
         pA == 'pA+' | pA == 'pA-',
         type != 'other', type != 'snRNA', type != 'XUT') %>%
  ungroup %>%
  mutate(pA = factor(pA, levels=c('pA+', 'pA+ + pA-', 'pA-', 'pA-X'))) %>%
ggplot(., aes(x=factor(1), y=total_Spnorm_signal_percent, fill=part)) + 
  geom_bar(stat='identity') +
  coord_polar(theta='y') +
  facet_grid(strain+type~pA+fraction+rapa+rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_blank())
```

## Violin plots per type and part


#### violin plot fun
```{r}
violin_plot <- function(df) {
  ggplot(df, aes(x=type, y=scaled_reads, fill=type)) +
  geom_violin(scale='width') +
  geom_boxplot(fill='gray', width=.2, outlier.shape = NA) +
  facet_grid(strain+part~pA+fraction+rep, scales='free') +
  scale_y_log10() +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45))
}
```


```{r violins all at once}
violin_plot(both_df)
```


```{r violins rapa0 rep1}
both_df %>%
  filter(rep == 1, rapa == 0, type != 'other') %>%
  violin_plot(.)
```

```{r violins rapa0 rep1 for Figure}
both_df %>%
  filter(rapa == 0, rep == 1, 
         pA == 'pA+' | pA == 'pA-',
         type != 'other', type != 'XUT', type != 'snRNA') %>%
  ungroup %>%
  mutate(pA = factor(pA, levels=c('pA+', 'pA+ + pA-', 'pA-', 'pA-X'))) %>%
  violin_plot(.)
```


```{r}
sessionInfo()
```



