---
title: "Correlation with published transcription studies"
author: "Manfred Schmid"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
`r format(Sys.time(), "%d %B, %Y")`


## SETUP
```{r global_options}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs_correlation_2min_with_published_txn/', dev='pdf',
                      echo=TRUE, warning=FALSE, message=FALSE)
```



```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(knitr))
suppressWarnings(library(broom))
```


## USING AVERAGED DATA 

### load the transcription estimates

from Markdown *Published_txn_measures.Rmd*
```{r}
load(file='../../data/published_gene_body_txn_estimates.RData', verbose=TRUE)

(txn_estimates %<>% separate(name, c('id', 'type', 'name', 'common_name'), sep=':'))
```





## Loading averaged  data

#### Load

avg over replicates from *2min_body_end_counts_to_R.Rmd*.  
```{r}
load('../../data/norm_2min_body_end_scaled_and_BGsub_avg_signals.RData', verbose = TRUE)

avg_signal
```

#### select only rapa 0 and pAplus or pAminus

```{r}
avg_signal %<>%
  ungroup %>%
  filter(rapa == 0,
         pA == 'pA+' | pA == 'pA-') %>%
  dplyr::select(-rapa)
```



#### clean up, combine and focus on 0 rapa samples

--> remove all genes not above BG in our data
```{r}
(both_df <- avg_signal %>%
   tidyr::unite(measure, c(part,pA, fraction), sep='_') %>%
   mutate(measure = sub('-', 'minus', measure) %>%
            sub('\\+', 'plus', .)) %>%
   spread(measure, mean_C) %>%
  left_join(., txn_estimates))
```




## correlation matrices

correlations are from data preformatted to matrix for simplicity
```{r}
mat <- both_df %>%
  ungroup %>%
  dplyr::select(-id, -name, -common_name, -size, 
                -contains('body_pAplus'),
                -contains('end_pAminus'))

head(mat)
```


#### Functions for plotting

```{r}
cor_df <- function(mat, ...) {
  
  cor_df <- cor(mat, ...) %>%
  as.data.frame() %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1)
  
  spearman_dists <- cor_dist(mat, ...)

  hc <- hclust(spearman_dists)

  hc_order <- hc$labels[hc$order]
  
  cor_df %<>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order))
}
```


```{r correlation matrix plotting}
cor_df_plot <- function(mat, study_order, ...) {
  
  mat %>%
    na.omit %>%
    cor(., ...) %>%
    as.data.frame() %>%
    rownames_to_column(var='study1') %>%
    gather(study2, cor, -study1) %>%
    mutate(study1 = factor(study1, levels=study_order),
           study2 = factor(study2, levels=study_order)) %>%
  ggplot(.,aes(x=study1, y=study2, fill=cor)) + 
  geom_tile() +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 6),
        axis.text.y=element_text(size = 6)) +
  geom_text(aes(label=round(cor,2)), size=2)
}
```


```{r correlation matrix p_values functions}
sig_level_class <- function(p) {
  ifelse(p < .00001, '*****',
   ifelse(p < .0001, '****',
    ifelse(p < .001, '***',
     ifelse(p < .01, '**',
      ifelse(p < .05, '*', 'ns')))))
}


cor_pvals <- function(mat, ...) {
  
  mat %>%
    na.omit %>%
    cor(., ...)

  cols <- expand.grid(colnames(mat), colnames(mat))
  
  (cor.pvalues <- apply(cols, 1, 
                     function(col) 
                       c(col[1], 
                         col[2], 
                         cor.test(mat[,col[1]],
                                  mat[,col[2]],
                                  ...)$p.value)))

  cor.pvalues <- data.frame(t(cor.pvalues), 
                            stringsAsFactors = FALSE)

  colnames(cor.pvalues) <- c('study1', 'study2', 'cor_pvalue')

  cor.pvalues$cor_pvalue <- as.numeric(cor.pvalues$cor_pvalue)

  cor.pvalues$sig_level <- sig_level_class(cor.pvalues$cor_pvalue)
  
  cor.pvalues$mlog10_cor_pvalue <- -log10(cor.pvalues$cor_pvalue)
  
  cor.pvalues$mlog10_cor_pvalue <- case_when(cor.pvalues$mlog10_cor_pvalue > 99 ~ '>99', 
                                             cor.pvalues$mlog10_cor_pvalue < 2 ~
                                               as.character(round(cor.pvalues$mlog10_cor_pvalue,2)),
                                             cor.pvalues$mlog10_cor_pvalue < 10 ~
                                               as.character(round(cor.pvalues$mlog10_cor_pvalue,1)),
                                             TRUE ~
                                               as.character(round(cor.pvalues$mlog10_cor_pvalue,0)))

  cor.pvalues
}

cor_pval_plot <- function(mat, study_order, ...) {
  
  cor.pvalues <- cor_pvals(mat, ...)

  cor.pvalues %>%
  mutate(study1 = factor(study1, levels=study_order),
         study2 = factor(study2, levels=study_order),
         sig_level = factor(sig_level,
                            levels = c('ns', '*', 
                                       '**', '***', 
                                       '****', '*****'))) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=sig_level)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  labs(fill="-log10(p)") +
  #scale_fill_gradientn(colors = c("white", "green4", 'darkgreen'),
  #                     values = scales::rescale(c(0,50,Inf))) +
  scale_fill_brewer(palette='OrRd', drop=FALSE) +
  geom_text(aes(label=mlog10_cor_pvalue))
}
```


```{r correlation hclust function}
cor_dist <- function(x, ...) {
  as.dist((1 - cor(x, ...))/2)
}

cor_dendro_hc <- function(mat, ...){
  mat %>%
    na.omit %>%
    cor_dist(., ...) %>%
    hclust
}
```


```{r scatter plotting function}
scatter_plots <- function(mat, xcol_name = 'NETseq_mean', ...) {
  
  cnames <- colnames(mat)
  x_col <- which(cnames == xcol_name)
  xvals <- mat[,x_col]
  
  par(mfrow=c(2, 4))
  for(y in 1:8) {
    yvals <- mat[,y]
    plot(xvals, yvals, 
         cex=.1, alpha=.2, pch=19,
         xlab=xcol_name,
         ylab=cnames[y], ...)
    valid_rows <- ( is.finite(xvals) & is.finite(yvals) )
    abline(lm(yvals[valid_rows]~xvals[valid_rows]), col='orange')
  }

  par(mfrow=c(1,1))
}
```




#### all at once

```{r}
log2_mat <- mat %>%
    dplyr::select(-type) %>%
    as.matrix %>%
    log2
```

```{r}
hc <- cor_dendro_hc(log2_mat, method='spearman', use='pairwise.complete.obs')
```

```{r dendrogram matrix all}
plot(hc)
```

```{r spearman correlation matrix all Figure}
cor_df_plot(log2_mat, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r spearman correlation pvalues matrix all Figure}
cor_pval_plot(log2_mat, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r scatter 3seq rel NETseq all Figure}
scatter_plots(log2_mat, xcol_name = 'NETseq_mean', 
              main = 'all, rel NETseq')
```





#### ORFs

```{r}
orf_rows <- (mat$type == 'ORF-T')

log2_mat_orfs <- log2_mat[orf_rows,]
```

```{r}
nrow(log2_mat_orfs)
```

```{r}
nrow(na.omit(log2_mat_orfs))
```

```{r}
hc <- cor_dendro_hc(log2_mat_orfs, method='spearman', use='pairwise.complete.obs')
```

```{r dendrogram matrix ORFs}
plot(hc)
```

```{r spearman correlation matrix ORFs Figure}
cor_df_plot(log2_mat_orfs, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r spearman correlation pvalues matrix ORFs Figure}
cor_pval_plot(log2_mat_orfs, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r scatter 3seq rel NETseq ORFs Figure}
scatter_plots(log2_mat_orfs, xcol_name = 'NETseq_mean', 
              main = 'ORFs, rel NETseq')
```



#### SUTs

```{r}
sut_rows <- (mat$type == 'SUTs')

log2_mat_suts <- log2_mat[sut_rows,]
```

```{r}
nrow(log2_mat_suts)
```

```{r}
nrow(na.omit(log2_mat_suts))
```

```{r}
hc <- cor_dendro_hc(log2_mat_suts, method='spearman', use='pairwise.complete.obs')
```

```{r dendrogram matrix SUTs}
plot(hc)
```

```{r spearman correlation matrix SUTs Figure}
cor_df_plot(log2_mat_suts, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r spearman correlation pvalues matrix SUTs Figure}
cor_pval_plot(log2_mat_suts, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r scatter 3seq rel NETseq SUTs Figure}
scatter_plots(log2_mat_suts, xcol_name = 'NETseq_mean', 
              main = 'SUTs, rel NETseq')
```



#### CUTs

```{r}
cut_rows <- (mat$type == 'CUTs')

log2_mat_cuts <- log2_mat[cut_rows,]
```

```{r}
nrow(log2_mat_cuts)
```

```{r}
nrow(na.omit(log2_mat_cuts))
```

```{r}
hc <- cor_dendro_hc(log2_mat_cuts, method='spearman', use='pairwise.complete.obs')
```

```{r dendrogram matrix CUTs}
plot(hc)
```

```{r spearman correlation matrix CUTs Figure}
cor_df_plot(log2_mat_cuts, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r spearman correlation pvalues matrix CUTs Figure}
cor_pval_plot(log2_mat_cuts, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r scatter 3seq rel NETseq CUTs Figure}
scatter_plots(log2_mat_cuts, xcol_name = 'NETseq_mean', 
              main = 'CUTs, rel NETseq')
```




## USING REPLICATES separately

-> from script body_end_counting.Rmd  
-> her we only look at length-normalized BGsub data
```{r}
load('../../data/norm_2min_body_end_scaled_and_BGsub.RData', verbose = TRUE)

both_df
```


#### clean up, combine and focus on 0 rapa samples

--> remove all genes not above BG in our data


```{r}
(both_df2 <- both_df %>%
   filter(rapa == 0, 
          pA == 'pA+' | pA == 'pA-') %>%
   tidyr::unite(measure, c(part, pA, strain, fraction, rep), sep='_') %>%
   dplyr::select(-Pap, -rapa, -sum) %>%
   mutate(measure = sub('-', 'minus', measure) %>%
            sub('\\+', 'plus', .)) %>%
   spread(measure, scaled_reads) %>%
   left_join(., txn_estimates))
```


```{r}
mat2 <- both_df2 %>%
  ungroup %>%
  dplyr::select(-id, -name, -common_name, -size, 
                -contains('body_pAplus'),
                -contains('end_pAminus'))

head(mat2)
```


#### all at once

```{r}
log2_mat <- mat2 %>%
    dplyr::select(-type) %>%
    as.matrix %>%
    log2
```

```{r}
hc <- cor_dendro_hc(log2_mat, method='spearman', use='pairwise.complete.obs')
```

```{r replicates dendrogram matrix all}
plot(hc)
```

```{r replicates spearman correlation matrix all Figure}
cor_df_plot(log2_mat, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r replicates spearman correlation pvalues matrix all Figure}
cor_pval_plot(log2_mat, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```


#### ORFs

```{r}
orf_rows <- (mat2$type == 'ORF-T')

log2_mat_orfs <- log2_mat[orf_rows,]
```

```{r}
hc <- cor_dendro_hc(log2_mat_orfs, method='spearman', use='pairwise.complete.obs')
```

```{r replicates dendrogram matrix ORFs}
plot(hc)
```

```{r replicates spearman correlation matrix ORFs Figure}
cor_df_plot(log2_mat_orfs, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r replicates spearman correlation pvalues matrix ORFs Figure}
cor_pval_plot(log2_mat_orfs, study_order = hc$labels[hc$order], 
            method='spearman', use='pairwise.complete.obs')
```

```{r replicates scatter 3seq rel NETseq ORFs Figure}
scatter_plots(log2_mat_orfs, xcol_name = 'NETseq_mean', 
              main = 'ORFs, rel NETseq')
```


```{r}
sessionInfo()
```

