---
title: 'Estimation of Decay rates 2 min data average of strains'
author: "Manfred Schmid"
output: 
  pdf_document: 
    toc: true 
    toc_depth: 3
    fig_caption: true
---


## Setup

`r format(Sys.time(), "%d %B, %Y")`

```{r setup, echo=TRUE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, 
                      fig.path=paste0('Figures_DR_2min/'), 
                      dev='pdf', 
                      echo=TRUE, warning=FALSE, message=FALSE, 
                      error=TRUE)
```



```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(DESeq2))
suppressWarnings(library(broom))
suppressWarnings(library(knitr))
```


## Loading triplicate raw data

#### Load

avg over replicates from *2min_body_end_counts_to_R.Rmd*.  
```{r}
load('../../data/norm_2min_body_end_scaled_and_BGsub_avg_signals.RData', verbose = TRUE)

avg_signal
```

#### select only rapa 0 and pAplus or pAminus

```{r}
avg_signal %<>%
  filter(rapa == 0,
         pA == 'pA+' | pA == 'pA-')
```


## DR from ip relative input ratio

#### Reorganize data input vs ip side-by-side

We can then in principle calculate the decay rate ... however there is still a major issue that the $sf$ from labelled and total RNA are not directly comparable. But estimation of the decay rate requires this. 

For a first analysis we assume they are directly comparable.


```{r}
(DR_ip_rel_in <- avg_signal %>%
  spread(fraction, mean_C) %>%
  mutate(ip_in_ratio = ip/input,
         DR = -log(1-ip_in_ratio)/2))
```


Check out the $ip/input$ ratios:  

```{r histogram of ip rel in ratio}
DR_ip_rel_in %>%
  ggplot(., aes(x=log10(ip_in_ratio), fill=part)) +
  geom_histogram() +
  facet_grid(part~pA)
```

We observe that $ip/total$ ratio is for the pA+ data clearly $<1$, which is OK. For the pA- data it is centered around 0, which would indicate full albelling, ie on average, especially for gene bodies, all RNA is labelled. This fits with the assumption that pAminus is often txn!


### IP rel in decay rate per type
```{r}
DR_ip_rel_in %<>%
  ungroup %>%
  mutate(type = factor(type, levels=c('ORF-T', 'SUTs', 'CUTs', 'other')))
```


#### save this
```{r}
save(DR_ip_rel_in, file='../../data/2min_DR_strainavg.RData')
```


#### DR between types

```{r DR violins part and pA facets}
DR_ip_rel_in %>%
  filter(type != 'other') %>%
  ggplot(., aes(x=type, y=DR, fill=type)) +
  geom_violin() +
  geom_boxplot(width=.2, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  facet_grid(part~pA) +
  scale_fill_brewer(palette='Set1') +
  theme_bw()
```


median DRs
```{r}
DR_ip_rel_in %>%
  filter(part == 'end', !is.na(DR), DR > 0) %>%
  group_by(part, pA, type) %>%
  summarize(median_DR = median(DR)) %>%
  spread(pA, median_DR)
```

#### add sn(o)RNA data

```{r}
load('../../data/2min_DR_snRNAs.RData', verbose=TRUE)

DR_snRNAs
```

we consider only the exact end position
```{r}
DR_snRNAs %<>% 
  ungroup %>%
  filter(rel_pos == 0, pA == 'pA+' | pA == 'pA-') %>% 
  mutate(part = 'end', type = 'snRNA') %>%
  dplyr::select(-rel_pos)
```

```{r}
DR_ip_rel_in_with_snRNA <- bind_rows(DR_snRNAs, DR_ip_rel_in) %>%
  mutate(type = factor(type, levels=c('ORF-T', 'SUTs', 'CUTs', 'snRNA', 'other')),
         pA = factor(pA, levels=c('pA+', 'pA-')))
```


```{r DR violins part and pA facets ends with snRNA pos 0 with jitter}
DR_ip_rel_in_with_snRNA %>%
  filter(type != 'other', part == 'end') %>%
  ggplot(., aes(x=type, y=DR, fill=type)) +
  geom_violin() +
  geom_boxplot(width=.2, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  geom_jitter(size=.1, alpha=.1, color='lightgray') +
  scale_y_log10() +
  facet_grid(part~pA) +
  scale_fill_brewer(palette='Set1') +
  theme_bw()
```

```{r DR violins part and pA facets ends with snRNA pos 0 wo jitter}
DR_ip_rel_in_with_snRNA %>%
  filter(type != 'other', part == 'end') %>%
  ggplot(., aes(x=type, y=DR, fill=type)) +
  geom_violin() +
  geom_boxplot(width=.2, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  facet_grid(part~pA) +
  scale_fill_brewer(palette='Set1') +
  theme_bw()
```

#### n for the violins
```{r}
DR_ip_rel_in_with_snRNA %>%
  filter(type != 'other', part == 'end', DR > 0) %>% #note DR>0 for others not shown on violin plot!!
  group_by(type, part, pA) %>%
  summarize(cnt = n())
```


### Comparison with published estimates

#### Load published decay rate datasets
from Markdown Published_decay_measures.Rmd

```{r}
load(file='../../data/published_decay_estimates.RData', verbose=TRUE)

DR_literature
```



```{r}
(DR_ends <- DR_ip_rel_in %>% 
  filter(part == 'end') %>%
  mutate(pA = ifelse(pA == 'pA+', 'pAplus', 'pAminus'),
         comparison = paste0('DR_', pA)) %>%
  dplyr::select(comparison, id, name, type, DR) %>%
  tidyr::spread(comparison, DR))
```


```{r}
(DR_ends %<>% left_join(., DR_literature))
```


#### correlation matrix

```{r}
cor_dist <- function(x, ...) {
  as.dist((1 - cor(x, ...))/2)
}
```

```{r}
DR_mat <- DR_ends %>%
  filter(type == 'ORF-T') %>%
  dplyr::select(-id, -name, -type) %>%
  as.matrix %>%
  na.omit

head(DR_mat)
```

#### spearman correlation

```{r}
(cor_df <- DR_mat %>%
  cor(., method='spearman') %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df)
```


```{r spearman correlation hclust}
spearman_dists <- DR_mat %>%
  cor_dist(., method='spearman')

hc <- hclust(spearman_dists)

plot(hc)

hc_order <- hc$labels[hc$order]
```


```{r spearman correlation matrix}
cor_df %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```



## DR from txn vs total RNA ratio


Idea here is that in a simple first order decay rate scenario, the decay rate can be directly derived from total RNA amount and transcription rate from the formula:

$$ RNA_{total} = \frac{k_{txn}}{k_{decay}} $$

and hence:  
$$ k_{decay} = \frac{k_{txn}}{RNA_{total}} $$


In our dataset $k_{txn}$ is proportional to $pA^-_{4sU, genebody}$  
and
$RNA_{total}$ is proportional to $pA^+_{input, geneend}$.


Do the math:

```{r}
(txn <- avg_signal %>%
   ungroup %>%
  filter(part == 'body' & pA == 'pA-' & fraction == 'ip') %>%
   dplyr::select(id, type, name, mean_C) %>%
   dplyr::rename(txn = mean_C))

(total <- avg_signal %>%
   ungroup %>%
  filter(part == 'end' & pA == 'pA+' & fraction == 'input') %>%
   dplyr::select(id, type, name, mean_C) %>%
   dplyr::rename(total = mean_C))


```

```{r}
(DR_txn_rel_total <- full_join(total, txn) %>%
  mutate(DR_txn_rel_total = txn/total))
```


```{r}
DR_txn_rel_total %<>%
  ungroup %>%
  mutate(type = factor(type, levels=c('ORF-T', 'SUTs', 'CUTs', 'other')))
```


#### save this
```{r}
save(DR_txn_rel_total, file='../../data/2min_DR_strainavg_txn_rel_total.RData')
```


#### DR between types

```{r DR txn rel total violins}
DR_txn_rel_total %>%
  filter(type != 'other') %>%
  ggplot(., aes(x=type, y=DR_txn_rel_total, fill=type)) +
  geom_violin() +
  geom_boxplot(width=.2, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  scale_fill_brewer(palette='Set1') +
  theme_bw() +
  coord_cartesian(ylim=c(.0001,1000))
```

n for those violins
```{r}
DR_txn_rel_total %>%
  filter(type != 'other', DR_txn_rel_total > 0) %>% #note DR>0 for others not shown on violin plot!!
  group_by(type) %>%
  summarize(cnt = n())
```


### Comparison with published estimates


```{r}
(DR_txn_rel_total %<>% left_join(., DR_literature))
```


#### correlation matrix

```{r}
cor_dist <- function(x, ...) {
  as.dist((1 - cor(x, ...))/2)
}
```

```{r}
DR_mat <- DR_txn_rel_total %>%
  filter(type == 'ORF-T') %>%
  dplyr::select(-id, -name, -type, -txn, -total) %>%
  as.matrix %>%
  na.omit

head(DR_mat)
```


n genes considered:
```{r}
nrow(DR_mat)
```

#### spearman correlation

```{r}
(cor_df <- DR_mat %>%
  cor(., method='spearman') %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df)
```


```{r spearman correlation hclust DR txn rel total}
spearman_dists <- DR_mat %>%
  cor_dist(., method='spearman')

hc <- hclust(spearman_dists)

plot(hc)

hc_order <- hc$labels[hc$order]
```


```{r spearman correlation matrix DR txn rel total}
cor_df %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```



## DR from body vs end in total RNA 


Idea is as before except that transcription is inferred from pAminus data from total instead of 4sU RNA.
Do the math:

```{r}
(txn_total <- avg_signal %>%
   ungroup %>%
  filter(part == 'body' & pA == 'pA-' & fraction == 'input') %>%
   dplyr::select(id, type, name, mean_C) %>%
   dplyr::rename(txn_from_total = mean_C))
```

```{r}
(DR_txn_rel_total %<>% full_join(., txn_total) %>%
  mutate(DR_txnfromtotal_rel_total = txn_from_total/total))
```


```{r}
DR_txn_rel_total %<>%
  ungroup %>%
  mutate(type = factor(type, levels=c('ORF-T', 'SUTs', 'CUTs', 'other')))
```


#### save this
```{r}
save(DR_txn_rel_total, file='../../data/2min_DR_strainavg_txn_rel_total.RData')
```


#### DR between types

```{r DR txn from total rel total violins}
DR_txn_rel_total %>%
  filter(type != 'other') %>%
  ggplot(., aes(x=type, y=DR_txnfromtotal_rel_total, fill=type)) +
  geom_violin() +
  geom_boxplot(width=.2, 
               outlier.shape = NA, 
               outlier.color = NULL, 
               fill='lightgray') +
  scale_y_log10() +
  scale_fill_brewer(palette='Set1') +
  theme_bw()+
  coord_cartesian(ylim=c(.0001,1000))
```

n for those violins
```{r}
DR_txn_rel_total %>%
  dplyr::select(id, type, DR_txnfromtotal_rel_total) %>%
  filter(type != 'other', DR_txnfromtotal_rel_total > 0) %>% #note DR>0 for others not shown on violin plot!!
  group_by(type) %>%
  summarize(cnt = n())
```



### Comparison with published estimates


```{r}
(DR_txn_rel_total %<>% left_join(., DR_literature))
```


#### correlation matrix

```{r}
DR_mat <- DR_txn_rel_total %>%
  filter(type == 'ORF-T') %>%
  dplyr::select(-id, -name, -type, -txn, -total, -txn_from_total) %>%
  as.matrix %>%
  na.omit

head(DR_mat)
```

n for correlations
```{r}
nrow(DR_mat)
```

#### spearman correlation

```{r}
(cor_df <- DR_mat %>%
  cor(., method='spearman') %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df)
```


```{r spearman correlation hclust DR txn from total rel total}
spearman_dists <- DR_mat %>%
  cor_dist(., method='spearman')

hc <- hclust(spearman_dists)

plot(hc)

hc_order <- hc$labels[hc$order]
```


```{r spearman correlation matrix DR txn from total rel total}
cor_df %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```



## compare all combined

```{r}
(DR_all <- full_join(DR_ends, DR_txn_rel_total))
```



```{r}
DR_mat <- DR_all %>%
  filter(type == 'ORF-T') %>%
  dplyr::select(-id, -name, -type, -txn, -total, -txn_from_total) %>%
  as.matrix %>%
  na.omit

head(DR_mat)
```


#### spearman correlation

```{r}
(cor_df <- DR_mat %>%
  cor(., method='spearman') %>%
  data.frame %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  tbl_df)
```

```{r}
cols <- expand.grid(colnames(DR_mat), colnames(DR_mat))
(cor.pvalues <- apply(cols, 1, 
                     function(col) 
                       c(col[1], 
                         col[2], 
                         cor.test(DR_mat[,col[1]],
                                  DR_mat[,col[2]],
                                  method='spearman')$p.value)))

cor.pvalues <- data.frame(t(cor.pvalues), stringsAsFactors = FALSE)
colnames(cor.pvalues) <- c('study1', 'study2', 'cor_pvalue')
cor.pvalues$cor_pvalue <- as.numeric(cor.pvalues$cor_pvalue)

cor.pvalues$mlog10_cor_pvalue <- -log10(cor.pvalues$cor_pvalue)

```


```{r spearman correlation hclust DR all}
spearman_dists <- DR_mat %>%
  cor_dist(., method='spearman')

hc <- hclust(spearman_dists)

plot(hc)

hc_order <- hc$labels[hc$order]
```


```{r spearman correlation matrix DR all}
cor_df %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```

```{r spearman pvalues matrix DR all}
sig_level_class <- function(p) {
  ifelse(p < .00001, '*****',
   ifelse(p < .0001, '****',
    ifelse(p < .001, '***',
     ifelse(p < .01, '**',
      ifelse(p < .05, '*', 'ns')))))
}

cor.pvalues$sig_level <- sig_level_class(cor.pvalues$cor_pvalue)
cor.pvalues$mlog10_cor_pvalue <- ifelse(cor.pvalues$mlog10_cor_pvalue > 99, 
                                        '>99',
                                        as.character(round(cor.pvalues$mlog10_cor_pvalue,0)))
  
cor.pvalues %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order),
         sig_level = factor(sig_level,
                            levels = c('ns', '*', 
                                       '**', '***', 
                                       '****', '*****'))) %>%
  ggplot(., 
       aes(x=study1, y=study2, fill=sig_level)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  labs(fill="-log10(p)") +
  scale_fill_brewer(palette='OrRd', drop=FALSE) +
  geom_text(aes(label=mlog10_cor_pvalue))
```


#### correlation matrix for paper
```{r}
paper_cols <- c('DR_pAplus','Miller_DR','Sun_DR','Presnyak_total_DR','DR_txn_rel_total','DR_txnfromtotal_rel_total')
```

n for correlations
```{r}
nrow(DR_mat)
```


```{r spearman correlation hclust DR paper sel}
spearman_dists <- DR_mat[,paper_cols] %>%
  cor_dist(., method='spearman')

hc <- hclust(spearman_dists)

plot(hc)

hc_order <- hc$labels[hc$order]
```

```{r spearman correlation matrix DR paper sel}
cor_df %>%
  filter(study1 %in% paper_cols & study2 %in% paper_cols) %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order)) %>%
ggplot(., 
       aes(x=study1, y=study2, fill=cor)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  geom_text(aes(label=round(cor,2)))
```

```{r spearman pvalues matrix DR paper sel}
sig_level_class <- function(p) {
  ifelse(p < .00001, '*****',
   ifelse(p < .0001, '****',
    ifelse(p < .001, '***',
     ifelse(p < .01, '**',
      ifelse(p < .05, '*', 'ns')))))
}

cor.pvalues$sig_level <- sig_level_class(cor.pvalues$cor_pvalue)
cor.pvalues$mlog10_cor_pvalue <- ifelse(cor.pvalues$mlog10_cor_pvalue > 99, 
                                        '>99',
                                        as.character(round(cor.pvalues$mlog10_cor_pvalue,0)))
  
cor.pvalues %>%
  filter(study1 %in% paper_cols & study2 %in% paper_cols) %>%
  mutate(study1 = factor(study1, levels=hc_order),
         study2 = factor(study2, levels=hc_order),
         sig_level = factor(sig_level,
                            levels = c('ns', '*', 
                                       '**', '***', 
                                       '****', '*****'))) %>%
  ggplot(., 
       aes(x=study1, y=study2, fill=sig_level)) +
  geom_tile() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  labs(fill="-log10(p)") +
  scale_fill_brewer(palette='OrRd', drop=FALSE) +
  geom_text(aes(label=mlog10_cor_pvalue))
```


```{r}
sessionInfo()
```
