---
title: "Correlation with published transcription studies"
author: "Manfred Schmid"
output: html_document
---
`r format(Sys.time(), "%d %B, %Y")`

```{r global_options}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs_correlation_2min_with_published_txn/', dev='pdf',
                      echo=TRUE, warning=FALSE, message=FALSE)
```



```{r load packages, echo=T, warning=F, message=F}
suppressWarnings(library(tidyverse))
suppressWarnings(library(magrittr))
suppressWarnings(library(knitr))
suppressWarnings(library(broom))
suppressWarnings(library(LSD))
```


### load the transcription estimates

from Markdown *Published_txn_measures.Rmd*
```{r}
load(file='data/published_gene_body_txn_estimates.RData', verbose=TRUE)

(txn_estimates %<>% separate(name, c('id', 'type', 'name', 'common_name'), sep=':'))
```



### load the 3' seq counts

-> from script body_end_counting.Rmd  
-> her we only look at length-normalized BGsub data
```{r}
load('data/norm_2min_body_end_scaled_and_BGsub.RData', verbose = TRUE)

both_df
```


### combine and focus on 0 rapa samples

--> remove all genes not above BG in our data
```{r}
(both_df %<>% 
   filter(rapa == 0, scaled_reads > 0) %>% 
   left_join(., txn_estimates))
```



## correlation matrices

correlations are from log-scaled data for simplicity
```{r}
our_data <- both_df %>%
  unite(sample, c(part,pA,strain,fraction,rapa,rep), sep='_') %>%
  dplyr::select(id, type, sample, scaled_reads) 

txn_data <- dplyr::select(txn_estimates, -name, -common_name, -size) %>%
  gather(sample, scaled_reads, -id, -type)

(log2_mat <- bind_rows(our_data, txn_data) %>%
    mutate(scaled_reads = log2(scaled_reads)) %>%
    spread(sample, scaled_reads))
```


#### Correlation matrix

```{r}
study_order <- colnames(log2_mat)[2:ncol(log2_mat)]

study_order2 <- study_order[order(sub('Mex67', 'Nab2', study_order) %>%
                                    sub('end_', 'aend', .) %>%
                                    sub('CRAC', 'nCRAC', .) %>%
                                    sub('Chip', 'oChIP', .))]
```

```{r}
cor_df <- function(mat, ...) {
  mat %>%
  as.matrix(.) %>%
  cor(., ...) %>%
  as.data.frame() %>%
  rownames_to_column(var='study1') %>%
  gather(study2, cor, -study1) %>%
  mutate(study1 = factor(study1, levels=study_order2),
         study2 = factor(study2, levels=study_order2))
}
```

```{r spearman correlation matrix all}
log2_mat %>%
  dplyr::select(-id, -type) %>%
  cor_df(., method='spearman', use='pairwise.complete.obs') %>%
  ggplot(.,aes(x=study1, y=study2, fill=cor)) + 
  geom_tile() +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 6),
        axis.text.y=element_text(size = 6)) 
```


```{r spearman correlation matrix ORFs Figure}
log2_mat %>%
  filter(type == 'ORF-T') %>%
  dplyr::select(-id, -type, -contains('pA+ + pA-'),
                -contains('body_pA+'), contains('mean'), -contains('pA-X'), -contains('end_pA-')) %>%
  na.omit %>%
  cor_df(., method='spearman', use='complete.obs') %>%
  ggplot(.,aes(x=study1, y=study2, fill=cor)) + 
  geom_tile() +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 6),
        axis.text.y=element_text(size = 6)) +
  geom_text(aes(label=round(cor,2)), size=2)
```

```{r spearman correlation matrix SUTs}
log2_mat %>%
  filter(type == 'SUTs') %>%
  dplyr::select(-id, -type, -contains('pA+ + pA-'),
                -contains('body_pA+'), contains('mean'), -contains('pA-X'), -contains('end_pA-')) %>%
  na.omit %>%
  cor_df(., method='spearman', use='complete.obs') %>%
  ggplot(.,aes(x=study1, y=study2, fill=cor)) + 
  geom_tile() +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 6),
        axis.text.y=element_text(size = 6)) +
  geom_text(aes(label=round(cor,2)), size=2)
```

```{r spearman correlation matrix CUTs}
log2_mat %>%
  filter(type == 'CUTs') %>%
  dplyr::select(-id, -type, -contains('pA+ + pA-'),
                -contains('body_pA+'), contains('mean'), -contains('pA-X'), -contains('end_pA-')) %>%
  na.omit %>%
  cor_df(., method='spearman', use='complete.obs') %>%
  ggplot(.,aes(x=study1, y=study2, fill=cor)) + 
  geom_tile() +
  scale_fill_gradient2(low='red', mid='white', high='blue', limits=c(-1,1)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 6),
        axis.text.y=element_text(size = 6)) +
  geom_text(aes(label=round(cor,2)), size=2)
```


## correlation scatters

```{r}
both_df %<>%
  group_by(part, pA, strain, fraction, rapa, rep)
```


#### main scatter plot fun
```{r}
heatscatter_plot <- function(df,
                             y_col = 'Chipseq_mean',
                             use_part = 'body', 
                             use_pA = 'pA-', 
                             use_strain = 'Mex67AA', 
                             use_fraction = 'ip', 
                             use_rapa = 0, 
                             use_rep = 1,
                             cor=T, 
                             colpal = 'bl2gr2rd', 
                             ...) {
  sel_df <- df %>%
  filter(part == use_part, pA == use_pA, strain == use_strain, fraction == use_fraction, rapa == use_rapa, rep == use_rep)

heatscatter(log2(sel_df$scaled_reads), log2(unlist(sel_df[, y_col])), xlab=paste0(use_part, ' ',use_pA, ': ',use_strain, '_',use_fraction, '_',use_rapa, '_', use_rep), ylab = y_col,
            cor=cor, colpal=colpal,...)

xcenter <- median(log2(sel_df$scaled_reads))
ycenter <- median(na.omit(log2(unlist(sel_df[, y_col]))))
intercept <- ycenter - xcenter
abline(a = intercept, b=1, lty=2)
abline(a = intercept+2, b=1, lty=3)
abline(a = intercept-2, b=1, lty=3)
}
```

```{r heatscatters vs Milligan CRAC}
heatscatter2x2 <- function(df, ycol = 'CRAC_mean', ...) {
  par(mfrow=c(2,2))
  heatscatter_plot(df, y_col = ycol, use_fraction = 'input', ...)

  heatscatter_plot(df, y_col = ycol, use_fraction = 'ip', ...)

  heatscatter_plot(df, y_col = ycol, use_part = 'end', use_pA = 'pA+', use_fraction = 'input', ...)

  heatscatter_plot(df, y_col = ycol, use_part = 'end', use_pA = 'pA+', use_fraction = 'ip', ...)
  par(mfrow=c(1,1))
}
```



#### with Rpb3 ChIP^2

Correlations
```{r}
both_df %>%
  summarise(cor_pearson_xmean = cor(scaled_reads, Chip2_mean, method='pearson', use='pairwise.complete'),
            cor_spearman_xmean = cor(scaled_reads, Chip2_mean, method='spearman', use='pairwise.complete')) %>%
  kable
```



```{r heatscatters body pA- end pA+ vs Rpb3 ChIP2 means pearson}
heatscatter2x2(both_df, ycol = 'Chip2_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIP2 means spearman}
heatscatter2x2(both_df, ycol = 'Chip2_mean', method='spearman')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIP2 means only ORFs spearman}
filter(both_df, type == 'ORF-T') %>%
heatscatter2x2(., ycol = 'Chip2_mean', method='spearman')
```


```{r heatscatters body pA- end pA+ vs Rpb3 ChIP2 means only SUTs spearman}
filter(both_df, type == 'SUTs') %>%
heatscatter2x2(., ycol = 'Chip2_mean', method='spearman')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIP2 means only CUTs}
filter(both_df, type == 'CUTs') %>%
heatscatter2x2(., ycol = 'Chip2_mean', method='pearson')
```


#### with Rpb3 ChIPseq

Correlations
```{r}
both_df %>%
  summarise(cor_pearson_xmean = cor(scaled_reads, Chipseq_mean, method='pearson', use='pairwise.complete'),
            cor_spearman_xmean = cor(scaled_reads, Chipseq_mean, method='spearman', use='pairwise.complete')) %>%
  kable
```



```{r heatscatters body pA- end pA+ vs Rpb3 ChIPseq means pearson}
#heatscatter2x2(both_df, ycol = 'Chipseq_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIPseq means spearman}
heatscatter2x2(both_df, ycol = 'Chipseq_mean', method='spearman')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIPseq means only ORFs}
filter(both_df, type == 'ORF-T') %>%
heatscatter2x2(., ycol = 'Chipseq_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIPseq means only SUTs}
filter(both_df, type == 'SUTs') %>%
heatscatter2x2(., ycol = 'Chipseq_mean', method='spearman')
```

```{r heatscatters body pA- end pA+ vs Rpb3 ChIPseq means only CUTs}
filter(both_df, type == 'CUTs') %>%
heatscatter2x2(., ycol = 'Chipseq_mean', method='spearman')
```


#### with NETseq

Correlations
```{r}
both_df %>%
  summarise(cor_pearson_xmean = cor(scaled_reads, NETseq_mean, method='pearson', use='pairwise.complete'),
            cor_spearman_xmean = cor(scaled_reads, NETseq_mean, method='spearman', use='pairwise.complete')) %>%
  kable
```



```{r heatscatters body pA- end pA+ vs NETseq}
heatscatter2x2(both_df, ycol = 'NETseq_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs NETseq only ORFs}
filter(both_df, type == 'ORF-T') %>%
heatscatter2x2(., ycol = 'NETseq_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs NETseq only SUTs}
filter(both_df, type == 'SUTs') %>%
heatscatter2x2(., ycol = 'NETseq_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs NETseq only CUTs}
filter(both_df, type == 'CUTs') %>%
heatscatter2x2(., ycol = 'NETseq_mean', method='pearson')
```



#### with RNAPII CRAC Milligan

Correlations
```{r}
both_df %>%
  summarise(cor_pearson_xmean = cor(scaled_reads, CRAC_mean, method='pearson', use='pairwise.complete'),
            cor_spearman_xmean = cor(scaled_reads, CRAC_mean, method='spearman', use='pairwise.complete')) %>%
  kable
```



```{r heatscatters body pA- end pA+ vs CRAC}
heatscatter2x2(both_df, ycol = 'CRAC_mean', method='pearson')
```


```{r heatscatters body pA- end pA+ vs CRAC only ORFs}
filter(both_df, type == 'ORF-T') %>%
heatscatter2x2(., ycol = 'CRAC_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs CRAC only SUTs}
filter(both_df, type == 'SUTs') %>%
heatscatter2x2(., ycol = 'CRAC_mean', method='pearson')
```

```{r heatscatters body pA- end pA+ vs CRAC only CUTs}
filter(both_df, type == 'CUTs') %>%
heatscatter2x2(., ycol = 'CRAC_mean', method='pearson')
```


```{r}
sessionInfo()
```

