---
title: "Create pAminus Tracks by Subtraction"
author: "Manfred Schmid"
date: "9 August 2017"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'Figures_pAminus_tracks/', dev='pdf')
```

```{r, message=FALSE,warning=FALSE}
suppressWarnings(library('tidyverse'))
suppressWarnings(library('magrittr'))
suppressWarnings(library('knitr'))
```


# pAminus = pA++pA- minus pA+

Using S.pombe normalized values: subtract values from pA+ libraries from EPAP libraries  
ups:
--> can result in negatives!!  
--> for IPs this can result in pA- values higher than pA++pA- since pA+ can be negative

uses the *subtract_bedgraph.py* python script as in BGSub.Rmd


apply this to all files
```{bash, eval=FALSE}
#!/bin/bash

#This script substracts the values at each positions covered in ie the noPap libraries from xPap libraries

## subtract all noPap input positions for filtering from xPap ##

cd /Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bedgraph/

## input files, each xPap has exactly one matching noPap ...
for f in norm_xPap_*_input_*.bedgraph
  do
    echo $f
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${f} ${f/xPap/noPap} \
      --chunk_size 100000 --value_col 4 -o ${f/xPap_/xPappAminus_}
  done


## ip files, each noPap has 3 matching xPap files, ie the replicates
## only do the BGsub samples
for f in norm_xPap_*_ip_*_KevinRoyAfiltered_BGsub.bedgraph
  do
    echo $f
    noPap_file=$(echo ${f/xPap/noPap} | sed s/_2_/_1_/g | sed s/_3_/_1_/g)
    echo ${noPap_file}
    python ~/ms_tools/MS_Metagene_Tools/subtract_bedgraph.py ${f} ${noPap_file} \
      --chunk_size 100000 --value_col 4 -o ${f/xPap_/xPappAminus_}
  done
```


#### get stats for the effect of pAminus vs pA+

counting signal and positions covered:  
```{bash, eval = FALSE}
cd /Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bedgraph/

#### total signals and positions above 0
echo "file sum_signal sum_positions" > pAminus.stats

for f in norm_noPap*input*.bedgraph
    do
        echo $f
        total=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' $f)
        total_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' $f)
        echo "${f/.bedgraph/} $total $total_pos" >> pAminus.stats
    done
    
for f in norm_xPap*input*.bedgraph
    do
        echo $f
        total=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' $f)
        total_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' $f)
        echo "${f/.bedgraph/} $total $total_pos" >> pAminus.stats
    done

for f in norm_*ip*BGsub.bedgraph
    do
        echo $f
        total=$(awk '{sum += $4*($3-$2)}END{printf("%8.2f",sum)}' $f)
        total_pos=$(awk '{if($4 > 0){sum += ($3-$2)}}END{printf("%8.2f",sum)}' $f)
        echo "${f/.bedgraph/} $total $total_pos" >> pAminus.stats
    done
```

```{r}
(stats_pAminus <- read.table('/Users/schmidm/Documents/Results/Lexogen_RNAseq_2/STAR_bw/norm_and_pA_filtered_bedgraph/pAminus.stats', header=TRUE) %>% 
   tbl_df %>%
  mutate(file = sub('^norm_', '', file) %>% 
           sub('_plus_KevinRoyAfiltered', '', .) %>%
           sub('_minus_KevinRoyAfiltered', '', .)) %>%
  group_by(file) %>%
  summarize(sum_signal = sum(sum_signal),
            sum_positions = sum(sum_positions)) %>%
  separate(file, c('Pap', 'strain', 'fraction', 'rapa', 'rep'), sep='_', extra='drop'))
```

```{r}
noPaps <- stats_pAminus %>% filter(Pap == 'noPap') %>% dplyr::select(-sum_positions, -rep) %>% spread(Pap, sum_signal)

Paps <- stats_pAminus %>% filter(Pap != 'noPap')%>% dplyr::select(-sum_positions) %>% spread(Pap, sum_signal)

(stats_pAminus_signal <- left_join(Paps, noPaps))
```

#### sanity check
```{r}
stats_pAminus_signal %>%
  mutate(pAminus_difference = xPappAminus - (xPap - noPap)) %>%
  arrange(pAminus_difference) %>%
  kable
```

--> only minor differences, probably rounding errors --> ignore


#### visualize signal
```{r}
(stats_signal <- stats_pAminus_signal %>%
  gather(Pap, sum_signal, -strain, -fraction, -rapa, -rep))
```

```{r}
filter(stats_signal, rapa == 0)
```

```{r pAplus minus and both signal}
filter(stats_signal, rapa == 0) %>%
  ggplot(., aes(x=rep, y=sum_signal, fill=Pap)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(fraction ~ strain, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw()
```


```{r}
(stats_signal_fractions <- filter(stats_signal, rapa == 0, Pap != 'xPap') %>%
  group_by(strain,fraction,rapa,rep) %>%
  mutate(frac_signal = sum_signal/sum(sum_signal),
         pA = ifelse(Pap == 'noPap', 'pA+', 'pA-')))
```

```{r piechart pAplus vs minus signal}
ggplot(stats_signal_fractions, aes(x=factor(1), y=frac_signal, fill=pA)) +
  geom_bar(stat='identity') +
  facet_grid(fraction ~ strain + rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  coord_polar(theta='y') +
  theme_bw()
```



#### positions covered comparison
```{r}
noPaps_pos <- stats_pAminus %>% filter(Pap == 'noPap') %>% dplyr::select(-sum_signal, -rep) %>% spread(Pap, sum_positions)

Paps_pos <- stats_pAminus %>% filter(Pap != 'noPap')%>% dplyr::select(-sum_signal) %>% spread(Pap, sum_positions)

(stats_pAminus_positions <- left_join(Paps_pos, noPaps_pos))
```



```{r}
(stats_positions <- stats_pAminus_positions %>%
  gather(Pap, sum_positions, -strain, -fraction, -rapa, -rep))
```

```{r}
filter(stats_positions, rapa == 0)
```

```{r pAplus minus and both positions}
filter(stats_positions, rapa == 0) %>%
  ggplot(., aes(x=rep, y=sum_positions, fill=Pap)) +
  geom_bar(stat='identity', position='dodge') +
  facet_grid(fraction ~ strain, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  theme_bw()
```


```{r}
(stats_positions_fractions <- filter(stats_positions, rapa == 0, Pap != 'xPap') %>%
  group_by(strain,fraction,rapa,rep) %>%
  mutate(frac_positions = sum_positions/sum(sum_positions),
         pA = ifelse(Pap == 'noPap', 'pA+', 'pA-')))
```

```{r piechart pAplus vs minus positions}
ggplot(stats_positions_fractions, aes(x=factor(1), y=frac_positions, fill=pA)) +
  geom_bar(stat='identity') +
  facet_grid(fraction ~ strain + rep, scales='free') +
  scale_fill_brewer(palette = 'Set1') +
  coord_polar(theta='y') +
  theme_bw()
```

####some sanity check:
```{bash, eval = FALSE}
###for SNR4
awk '$1=="chrV" && $2 == 424882' norm_noPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bedgraph
#chrV	424882	424883	-111.6035


awk '$1=="chrV" && $2 == 424882' norm_xPap_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bedgraph
#chrV	424882	424883	802.2


awk '$1=="chrV" && $2 == 424882' norm_xPappAminus_Mex67AA_ip_0_1_plus_KevinRoyAfiltered_BGsub.bedgraph
#chrV	424882	424883	913.8035

# --> ups this oddity since negatives are allowed in both BGsub files !!


awk '$1=="chrV" && $2 == 424882' norm_noPap_Mex67AA_input_0_1_plus_KevinRoyAfiltered.bedgraph
#chrV	424882	424883	249.642


awk '$1=="chrV" && $2 == 424882' norm_xPap_Mex67AA_input_0_1_plus_KevinRoyAfiltered.bedgraph
#chrV	424882	424883	362039


awk '$1=="chrV" && $2 == 424882' norm_xPappAminus_Mex67AA_input_0_1_plus_KevinRoyAfiltered.bedgraph
#chrV	424882	424883	361789.358
```

--> ups this oddity since negatives are allowed !!
--> otherwise seems to be OK..



```{r}
sessionInfo()
```



